### 2021-01-09 paper/ijcai21.tex
Start tracking this

%%%% ijcai21.tex

\typeout{IJCAI--21 Instructions for Authors}

% These are the instructions for authors for IJCAI-21.

\documentclass{article}
\pdfpagewidth=8.5in
\pdfpageheight=11in
% The file ijcai21.sty is NOT the same than previous years'
\usepackage{ijcai21}

% Use the postscript times font!
\usepackage{times}
\usepackage{soul}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmic}
\urlstyle{same}

% the following package is optional:
%\usepackage{latexsym}

% See https://www.overleaf.com/learn/latex/theorems_and_proofs
% for a nice explanation of how to define new theorems, but keep
% in mind that the amsthm package is already included in this
% template and that you must *not* alter the styling.
\newtheorem{example}{Example}
\newtheorem{theorem}{Theorem}

% Following comment is from ijcai97-submit.tex:
% The preparation of these files was supported by Schlumberger Palo Alto
% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
% preparation.

% These instructions can be modified and used in other conferences as long
% as credit to the authors and supporting agencies is retained, this notice
% is not changed, and further modification or reuse is not restricted.
% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
% contacts for providing assistance without their prior permission.

% To use for other conferences, change references to files and the
% conference appropriate and use other authors, contacts, publishers, and
% organizations.
% Also change the deadline and address for returning papers and the length and
% page charge instructions.
% Put where the files are available in the appropriate places.

%PDF Info Is REQUIRED.
\pdfinfo{
/TemplateVersion (IJCAI.2021.0)
}

\title{IJCAI--21 Formatting Instructions}

% Single author syntax
\author{
    Zhi-Hua Zhou
    \affiliations
    Nanjing University
    \emails
    pcchair@ijcai-21.org
}

% Multiple author syntax (remove the single-author syntax above and the \iffalse ... \fi here)
% Check the ijcai21-multiauthor.tex file for detailed instructions
\iffalse
\author{
First Author$^1$
\and
Second Author$^2$\and
Third Author$^{2,3}$\And
Fourth Author$^4$
\affiliations
$^1$First Affiliation\\
$^2$Second Affiliation\\
$^3$Third Affiliation\\
$^4$Fourth Affiliation
\emails
\{first, second\}@example.com,
third@other.example.com,
fourth@example.com
}
\fi

\begin{document}

\maketitle

\begin{abstract}
  The {\it IJCAI--21 Proceedings} will be printed from electronic
  manuscripts submitted by the authors. The electronic manuscript will
  also be included in the online version of the proceedings. This paper
  provides the style instructions.
\end{abstract}

\section{Introduction}

The {\it IJCAI--21 Proceedings} will be printed from electronic
manuscripts submitted by the authors. These must be PDF ({\em Portable
Document Format}) files formatted for 8-1/2$''$ $\times$ 11$''$ paper.

\subsection{Length of Papers}

All paper {\em submissions} must have a maximum of six pages, plus at most one for references. The seventh page cannot contain {\bf anything} other than references.

The length rules may change for final camera-ready versions of accepted papers and will differ between tracks. Some tracks may include only references in the last page, whereas others allow for any content in all pages. Similarly, some tracks allow you to buy a few extra pages should you want to, whereas others don't.

If your paper is accepted, please carefully read the notifications you receive, and check the proceedings submission information website\footnote{\url{https://proceedings.ijcai.org/info}} to know how many pages you can finally use. That website holds the most up-to-date information regarding paper length limits at all times. Please notice that if your track allows for a special references-only page, the {\bf references-only page(s) cannot contain anything else than references} (i.e.: do not write your acknowledgments on that page or you will be charged for it).

\subsection{Word Processing Software}

As detailed below, IJCAI has prepared and made available a set of
\LaTeX{} macros and a Microsoft Word template for use in formatting
your paper. If you are using some other word processing software, please follow the format instructions given below and ensure that your final paper looks as much like this sample as possible.

\section{Style and Format}

\LaTeX{} and Word style files that implement these instructions
can be retrieved electronically. (See Appendix~\ref{stylefiles} for
instructions on how to obtain these files.)

\subsection{Layout}

Print manuscripts two columns to a page, in the manner in which these
instructions are printed. The exact dimensions for pages are:
\begin{itemize}
\item left and right margins: .75$''$
\item column width: 3.375$''$
\item gap between columns: .25$''$
\item top margin---first page: 1.375$''$
\item top margin---other pages: .75$''$
\item bottom margin: 1.25$''$
\item column height---first page: 6.625$''$
\item column height---other pages: 9$''$
\end{itemize}

All measurements assume an 8-1/2$''$ $\times$ 11$''$ page size. For
A4-size paper, use the given top and left margins, column width,
height, and gap, and modify the bottom and right margins as necessary.

\subsection{Format of Electronic Manuscript}

For the production of the electronic manuscript, you must use Adobe's
{\em Portable Document Format} (PDF). A PDF file can be generated, for
instance, on Unix systems using {\tt ps2pdf} or on Windows systems
using Adobe's Distiller. There is also a website with free software
and conversion services: \url{http://www.ps2pdf.com}. For reasons of
uniformity, use of Adobe's {\em Times Roman} font is strongly suggested.
In \LaTeX2e{} this is accomplished by writing
\begin{quote}
\mbox{\tt $\backslash$usepackage\{times\}}
\end{quote}
in the preamble.\footnote{You may want also to use the package {\tt
latexsym}, which defines all symbols known from the old \LaTeX{}
version.}

Additionally, it is of utmost importance to specify the {\bf
letter} format (corresponding to 8-1/2$''$ $\times$ 11$''$) when
formatting the paper. When working with {\tt dvips}, for instance, one
should specify {\tt -t letter}.

\subsection{Title and Author Information}

Center the title on the entire width of the page in a 14-point bold
font. The title must be capitalized using Title Case. Below it, center author name(s) in 12-point bold font. On the following line(s) place the affiliations, each affiliation on its own line using 12-point regular font. Matching between authors and affiliations can be done using numeric superindices. Optionally, a comma-separated list of email addresses follows the affiliation(s) line(s), using 12-point regular font.

\subsubsection{Blind Review}

In order to make blind reviewing possible, authors must omit their
names and affiliations when submitting the paper for review. In place
of names and affiliations, provide a list of content areas. When
referring to one's own work, use the third person rather than the
first person. For example, say, ``Previously,
Gottlob~\shortcite{gottlob:nonmon} has shown that\ldots'', rather
than, ``In our previous work~\cite{gottlob:nonmon}, we have shown
that\ldots'' Try to avoid including any information in the body of the
paper or references that would identify the authors or their
institutions. Such information can be added to the final camera-ready
version for publication.

\subsection{Abstract}

Place the abstract at the beginning of the first column 3$''$ from the
top of the page, unless that does not leave enough room for the title
and author information. Use a slightly smaller width than in the body
of the paper. Head the abstract with ``Abstract'' centered above the
body of the abstract in a 12-point bold font. The body of the abstract
should be in the same font as the body of the paper.

The abstract should be a concise, one-paragraph summary describing the
general thesis and conclusion of your paper. A reader should be able
to learn the purpose of the paper and the reason for its importance
from the abstract. The abstract should be no more than 200 words long.

\subsection{Text}

The main body of the text immediately follows the abstract. Use
10-point type in a clear, readable font with 1-point leading (10 on
11).

Indent when starting a new paragraph, except after major headings.

\subsection{Headings and Sections}

When necessary, headings should be used to separate major sections of
your paper. (These instructions use many headings to demonstrate their
appearance; your paper should have fewer headings.). All headings should be capitalized using Title Case.

\subsubsection{Section Headings}

Print section headings in 12-point bold type in the style shown in
these instructions. Leave a blank space of approximately 10 points
above and 4 points below section headings.  Number sections with
arabic numerals.

\subsubsection{Subsection Headings}

Print subsection headings in 11-point bold type. Leave a blank space
of approximately 8 points above and 3 points below subsection
headings. Number subsections with the section number and the
subsection number (in arabic numerals) separated by a
period.

\subsubsection{Subsubsection Headings}

Print subsubsection headings in 10-point bold type. Leave a blank
space of approximately 6 points above subsubsection headings. Do not
number subsubsections.

\paragraph{Titled paragraphs.} You should use titled paragraphs if and
only if the title covers exactly one paragraph. Such paragraphs should be
separated from the preceding content by at least 3pt, and no more than
6pt. The title should be in 10pt bold font and ended with a period.
After that, a 1em horizontal space should follow the title before
the paragraph's text.

In \LaTeX{} titled paragraphs should be typeset using
\begin{quote}
{\tt \textbackslash{}paragraph\{Title.\} text} .
\end{quote}

\subsubsection{Acknowledgements}

You may include an unnumbered acknowledgments section, including
acknowledgments of help from colleagues, financial support, and
permission to publish. If present, acknowledgements must be in a dedicated,
unnumbered section appearing after all regular sections but before any
appendices or references.

Use
\begin{quote}
    {\tt \textbackslash{}section*\{Acknowledgements\}})
\end{quote}
to typeset the acknowledgements section in \LaTeX{}.

\subsubsection{Appendices}

Any appendices directly follow the text and look like sections, except
that they are numbered with capital letters instead of arabic
numerals. See this document for an example.

\subsubsection{References}

The references section is headed ``References'', printed in the same
style as a section heading but without a number. A sample list of
references is given at the end of these instructions. Use a consistent
format for references. The reference list should not include publicly unavailable work.

\subsection{Citations}

Citations within the text should include the author's last name and
the year of publication, for example~\cite{gottlob:nonmon}.  Append
lowercase letters to the year in cases of ambiguity.  Treat multiple
authors as in the following examples:~\cite{abelson-et-al:scheme}
or~\cite{bgf:Lixto} (for more than two authors) and
\cite{brachman-schmolze:kl-one} (for two authors).  If the author
portion of a citation is obvious, omit it, e.g.,
Nebel~\shortcite{nebel:jair-2000}.  Collapse multiple citations as
follows:~\cite{gls:hypertrees,levesque:functional-foundations}.
\nocite{abelson-et-al:scheme}
\nocite{bgf:Lixto}
\nocite{brachman-schmolze:kl-one}
\nocite{gottlob:nonmon}
\nocite{gls:hypertrees}
\nocite{levesque:functional-foundations}
\nocite{levesque:belief}
\nocite{nebel:jair-2000}

\subsection{Footnotes}

Place footnotes at the bottom of the page in a 9-point font.  Refer to
them with superscript numbers.\footnote{This is how your footnotes
should appear.} Separate them from the text by a short
line.\footnote{Note the line separating these footnotes from the
text.} Avoid footnotes as much as possible; they interrupt the flow of
the text.

\section{Illustrations}

Place all illustrations (figures, drawings, tables, and photographs)
throughout the paper at the places where they are first discussed,
rather than at the end of the paper.

They should be floated to the top (preferred) or bottom of the page,
unless they are an integral part
of your narrative flow. When placed at the bottom or top of
a page, illustrations may run across both columns, but not when they
appear inline.

Illustrations must be rendered electronically or scanned and placed
directly in your document. They should be cropped outside latex, otherwise portions of the image could reappear during the post-processing of your paper. All illustrations should be understandable when printed in black and
white, albeit you can use colors to enhance them. Line weights should
be 1/2-point or thicker. Avoid screens and superimposing type on
patterns, as these effects may not reproduce well.

Number illustrations sequentially. Use references of the following
form: Figure 1, Table 2, etc. Place illustration numbers and captions
under illustrations. Leave a margin of 1/4-inch around the area
covered by the illustration and caption.  Use 9-point type for
captions, labels, and other text in illustrations. Captions should always appear below the illustration.

\section{Tables}

Tables are considered illustrations containing data. Therefore, they should also appear floated to the top (preferably) or bottom of the page, and with the captions below them.

\begin{table}
\centering
\begin{tabular}{lll}
\hline
Scenario  & $\delta$ & Runtime \\
\hline
Paris       & 0.1s  & 13.65ms     \\
Paris       & 0.2s  & 0.01ms      \\
New York    & 0.1s  & 92.50ms     \\
Singapore   & 0.1s  & 33.33ms     \\
Singapore   & 0.2s  & 23.01ms     \\
\hline
\end{tabular}
\caption{Latex default table}
\label{tab:plain}
\end{table}

\begin{table}
\centering
\begin{tabular}{lrr}
\toprule
Scenario  & $\delta$ (s) & Runtime (ms) \\
\midrule
Paris       & 0.1  & 13.65      \\
            & 0.2  & 0.01       \\
New York    & 0.1  & 92.50      \\
Singapore   & 0.1  & 33.33      \\
            & 0.2  & 23.01      \\
\bottomrule
\end{tabular}
\caption{Booktabs table}
\label{tab:booktabs}
\end{table}

If you are using \LaTeX, you should use the {\tt booktabs} package, because it produces better tables than the standard ones. Compare Tables \ref{tab:plain} and~\ref{tab:booktabs}. The latter is clearly more readable for three reasons:

\begin{enumerate}
    \item The styling is better thanks to using the {\tt booktabs} rulers instead of the default ones.
    \item Numeric columns are right-aligned, making it easier to compare the numbers. Make sure to also right-align the corresponding headers, and to use the same precision for all numbers.
    \item We avoid unnecessary repetition, both between lines (no need to repeat the scenario name in this case) as well as in the content (units can be shown in the column header).
\end{enumerate}

\section{Formulas}

IJCAI's two-column format makes it difficult to typeset long formulas. A usual temptation is to reduce the size of the formula by using the {\tt small} or {\tt tiny} sizes. This doesn't work correctly with the current \LaTeX{} versions, breaking the line spacing of the preceding paragraphs and title, as well as the equation number sizes. The following equation demonstrates the effects (notice that this entire paragraph looks badly formatted):
%
\begin{tiny}
\begin{equation}
    x = \prod_{i=1}^n \sum_{j=1}^n j_i + \prod_{i=1}^n \sum_{j=1}^n i_j + \prod_{i=1}^n \sum_{j=1}^n j_i + \prod_{i=1}^n \sum_{j=1}^n i_j + \prod_{i=1}^n \sum_{j=1}^n j_i
\end{equation}
\end{tiny}%

Reducing formula sizes this way is strictly forbidden. We {\bf strongly} recommend authors to split formulas in multiple lines when they don't fit in a single line. This is the easiest approach to typeset those formulas and provides the most readable output%
%
\begin{align}
    x =& \prod_{i=1}^n \sum_{j=1}^n j_i + \prod_{i=1}^n \sum_{j=1}^n i_j + \prod_{i=1}^n \sum_{j=1}^n j_i + \prod_{i=1}^n \sum_{j=1}^n i_j + \nonumber\\
    + & \prod_{i=1}^n \sum_{j=1}^n j_i
\end{align}%

If a line is just slightly longer than the column width, you may use the {\tt resizebox} environment on that equation. The result looks better and doesn't interfere with the paragraph's line spacing: %
\begin{equation}
\resizebox{.91\linewidth}{!}{$
    \displaystyle
    x = \prod_{i=1}^n \sum_{j=1}^n j_i + \prod_{i=1}^n \sum_{j=1}^n i_j + \prod_{i=1}^n \sum_{j=1}^n j_i + \prod_{i=1}^n \sum_{j=1}^n i_j + \prod_{i=1}^n \sum_{j=1}^n j_i
$}
\end{equation}%

This last solution may have to be adapted if you use different equation environments, but it can generally be made to work. Please notice that in any case:

\begin{itemize}
    \item Equation numbers must be in the same font and size than the main text (10pt).
    \item Your formula's main symbols should not be smaller than {\small small} text (9pt).
\end{itemize}

For instance, the formula
%
\begin{equation}
    \resizebox{.91\linewidth}{!}{$
    \displaystyle
    x = \prod_{i=1}^n \sum_{j=1}^n j_i + \prod_{i=1}^n \sum_{j=1}^n i_j + \prod_{i=1}^n \sum_{j=1}^n j_i + \prod_{i=1}^n \sum_{j=1}^n i_j + \prod_{i=1}^n \sum_{j=1}^n j_i + \prod_{i=1}^n \sum_{j=1}^n i_j
$}
\end{equation}
%
would not be acceptable because the text is too small.

\section{Examples, Definitions, Theorems and Similar}

Examples, definitions, theorems, corollaries and similar must be written in their own paragraph. The paragraph must be separated by at least 2pt and no more than 5pt from the preceding and succeeding paragraphs. They must begin with the kind of item written in 10pt bold font followed by their number (e.g.: Theorem 1), optionally followed by a title/summary between parentheses in non-bold font and ended with a period. After that the main body of the item follows, written in 10 pt italics font (see below for examples).

In \LaTeX{} We strongly recommend you to define environments for your examples, definitions, propositions, lemmas, corollaries and similar. This can be done in your \LaTeX{} preamble using \texttt{\textbackslash{newtheorem}} -- see the source of this document for examples. Numbering for these items must be global, not per-section (e.g.: Theorem 1 instead of Theorem 6.1).

\begin{example}[How to write an example]
Examples should be written using the example environment defined in this template.
\end{example}

\begin{theorem}
This is an example of an untitled theorem.
\end{theorem}

You may also include a title or description using these environments as shown in the following theorem.

\begin{theorem}[A titled theorem]
This is an example of a titled theorem.
\end{theorem}

\section{Proofs}

Proofs must be written in their own paragraph separated by at least 2pt and no more than 5pt from the preceding and succeeding paragraphs. Proof paragraphs should start with the keyword ``Proof." in 10pt italics font. After that the proof follows in regular 10pt font. At the end of the proof, an unfilled square symbol (qed) marks the end of the proof.

In \LaTeX{} proofs should be typeset using the \texttt{\textbackslash{proof}} environment.

\begin{proof}
This paragraph is an example of how a proof looks like using the \texttt{\textbackslash{proof}} environment.
\end{proof}

\section{Algorithms and Listings}

Algorithms and listings are a special kind of figures. Like all illustrations, they should appear floated to the top (preferably) or bottom of the page. However, their caption should appear in the header, left-justified and enclosed between horizontal lines, as shown in Algorithm~\ref{alg:algorithm}. The algorithm body should be terminated with another horizontal line. It is up to the authors to decide whether to show line numbers or not, how to format comments, etc.

In \LaTeX{} algorithms may be typeset using the {\tt algorithm} and {\tt algorithmic} packages, but you can also use one of the many other packages for the task.

\begin{algorithm}[tb]
\caption{Example algorithm}
\label{alg:algorithm}
\textbf{Input}: Your algorithm's input\\
\textbf{Parameter}: Optional list of parameters\\
\textbf{Output}: Your algorithm's output
\begin{algorithmic}[1] %[1] enables line numbers
\STATE Let $t=0$.
\WHILE{condition}
\STATE Do some action.
\IF {conditional}
\STATE Perform task A.
\ELSE
\STATE Perform task B.
\ENDIF
\ENDWHILE
\STATE \textbf{return} solution
\end{algorithmic}
\end{algorithm}

\section*{Acknowledgments}

The preparation of these instructions and the \LaTeX{} and Bib\TeX{}
files that implement them was supported by Schlumberger Palo Alto
Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
Preparation of the Microsoft Word file was supported by IJCAI.  An
early version of this document was created by Shirley Jowell and Peter
F. Patel-Schneider.  It was subsequently modified by Jennifer
Ballentine and Thomas Dean, Bernhard Nebel, Daniel Pagenstecher,
Kurt Steinkraus, Toby Walsh and Carles Sierra. The current version
has been prepared by Marc Pujol-Gonzalez and Francisco Cruz-Mencia.

\appendix

\section{\LaTeX{} and Word Style Files}\label{stylefiles}

The \LaTeX{} and Word style files are available on the IJCAI--21
website, \url{https://www.ijcai21.org/}.
These style files implement the formatting instructions in this
document.

The \LaTeX{} files are {\tt ijcai21.sty} and {\tt ijcai21.tex}, and
the Bib\TeX{} files are {\tt named.bst} and {\tt ijcai21.bib}. The
\LaTeX{} style file is for version 2e of \LaTeX{}, and the Bib\TeX{}
style file is for version 0.99c of Bib\TeX{} ({\em not} version
0.98i). The {\tt ijcai21.sty} style differs from the {\tt
ijcai20.sty} file used for IJCAI--PRICAI--20.

The Microsoft Word style file consists of a single file, {\tt
ijcai21.doc}. This template differs from the one used for
IJCAI--PRICAI--20.

These Microsoft Word and \LaTeX{} files contain the source of the
present document and may serve as a formatting sample.

Further information on using these styles for the preparation of
papers for IJCAI--21 can be obtained by contacting {\tt
pcchair@ijcai-21.org}.

%% The file named.bst is a bibliography style file for BibTeX 0.99c
\bibliographystyle{named}
\bibliography{ijcai21}

\end{document}

### 2021-01-09 paper/paper.tex
Start tracking this

\documentclass{article}
\pdfpagewidth=8.5in
\pdfpageheight=11in
\usepackage{ijcai21}
\usepackage{times}
\usepackage{soul}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmic}
\urlstyle{same}

% \usepackage{showframe}

\newtheorem{proposition}{Proposition}

\usepackage{complexity}

\pdfinfo{
/TemplateVersion (IJCAI.2021.0)
}

\title{Solving Graph Homomorphism and Subgraph Isomorphism Problems Faster Through Neighbourhood
Maximum Clique Constraints}

\author{
Ciaran McCreesh$^1$
\and
Sonja Kraiczy$^2$
\affiliations
$^1$University of Glasgow\\
$^2$University of Oxford\\
\emails
ciaran.mccreesh@glasgow.ac.uk
}

\newcommand{\neighbourhood}{\operatorname{N}}
\newcommand{\vertexset}{\operatorname{V}}
\newcommand{\degree}{\operatorname{deg}}
\newcommand{\nds}{\operatorname{S}}

\begin{document}

\maketitle

\begin{abstract}
    We can find graph homomorphisms faster through clique constraints.
\end{abstract}

\section{Introduction}

A \emph{graph homomorphism} is a function between two graphs that maps adjacent vertices to adjacent
vertices. An injective graph homomorphism is more commonly known as a \emph{subgraph isomorphism},
whilst a \emph{locally injective homomorphism} is one which is injective when restricted to any
individual vertex and its neighbourhood \cite{DBLP:journals/csr/FialaK08}. Finding any of these three kinds of homomorphism between
two given graphs is \NP-complete.  The subgraph isomorphism problem arises in many practical
applications, including ??, and as such there has been a lot of research into finding algorithms
which perform well empirically. Most approaches are based either upon very fast but simple
backtracking algorithms,
\cite{DBLP:journals/pami/CordellaFSV04,DBLP:journals/bmcbi/BonniciGPSF13,DBLP:conf/gbrpr/CarlettiFSV17}
which often but not always perform well on very easy instances, or upon constraint programming
algorithms
\cite{DBLP:journals/constraints/ZampelliDS10,DBLP:journals/ai/Solnon10,DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15,DBLP:conf/cpaior/ArchibaldDHMP019},
which have higher startup costs but that perform vastly better on harder instances and much more
consistently on easy instances \cite{DBLP:journals/jair/McCreeshPST18,DBLP:conf/gbrpr/Solnon19}. The
current state of the art is the Glasgow Subgraph Solver \cite{DBLP:conf/gg/McCreeshP020}, which is a
dedicated constraint programming solver for subgraph-finding problems. Much of its performance comes
from inference strategies based upon degrees and neighbourhood degree sequences
\cite{DBLP:journals/constraints/ZampelliDS10}, counting paths between vertices
\cite{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}, and cardinality reasoning
\cite{DBLP:journals/ai/Solnon10}, which can be used to eliminate many infeasible candidate
assignments without search.

One might hope that this inference would also be helpful for other graph homomorphism problem
variants. Indeed, this paper proves that many, but not all, of these strategies are also valid in
the locally injective case, and that a weakened form of one other strategy is valid for all
homomorphisms. However, we also prove that \emph{none} of the other strategies are valid for finding
homomorphisms where there is no injectivity requirement. Finally, we introduce a new filtering
technique that is based upon finding the maximum clique in the neighbourhood of each vertex, and
prove that this technique is valid even in the general case. Although this new filtering technique
involves solving a large number of \NP-complete problems, we demonstrate that it is effective in
practice, particularly for the non-injective problem where filtering can solve instances ??many
hundreds of times faster in aggregate. This shows, for the first time, that modern constraint
programming techniques can be practical for less constrained graph homomorphism finding problems;
previous algorithmic approaches have focused instead upon worst-case computational bounds
\cite{DBLP:journals/csr/FialaK08,DBLP:journals/ipl/Rzazewski14,DBLP:journals/tcs/ChaplickFHPT15},
whose practical utility has yet to be demonstrated. The empirical effectiveness of our results is
especially important because many current applications use subgraph isomorphism solvers only because
they perform well using off the shelf solvers, rather than because they exactly match ideal domain
requirements ??cite.

\section{Background and Theory}

Let $G$ and $H$ be graphs. Let $v \in \vertexset(G)$ be a vertex of $G$. The \emph{(open)
neighbourhood} of $v$, written $\neighbourhood_G(v)$, is the set of vertices adjacent to $v$ not
including $v$ itself, whilst the closed neighbourhood of $v$, written $\neighbourhood_G[v]$, is the
neighbourhood plus $v$. The \emph{degree} of a vertex, $\degree_G(v)$, is the cardinality of its open
neighbourhood. Given a vertex set $S \subseteq \vertexset(G)$, the subgraph \emph{induced by} $S$,
written $G[S]$, is the subgraph of $G$ with only the vertices in $S$ together with all the edges
between them. A \emph{clique} is a subgraph where every vertex is adjacent to every other in the
subgraph.

A \emph{homomorphism} from $G$ to $H$ is a function mapping vertices of $G$ to vertices of $H$, such
that adjacent vertices in $G$ are mapped to adjacent vertices in $H$. A homomorphism $h$ is
\emph{locally injective} if for every vertex $w$, the restriction of $h$ to $G[N_G[w]]$ is
injective; if $h$ is injective globally we call it a \emph{(non-induced) subgraph isomorphism}.

?? loops, be really careful with this

?? CSP, hom as CSP

An \emph{implied constraint} is a constraint that can be added to a CSP without altering its set of
solutions. ?? Is it best wording all of this in terms of implied constraints?

\begin{proposition}For the problem of finding a subgraph isomorphism $i$ from a graph $G$ to a graph
    $H$, the following constraints are implied for any vertex $v \in \vertexset(G)$:
    \begin{enumerate}
        \item We cannot map $v$ to a vertex of lower degree, $\degree(v) \le \degree(i(v))$.
        \item Furthermore, ??NDS.
    \end{enumerate}
\end{proposition}

?? Show that these work for locally injective but not homomorphism

\begin{proposition}For the problem of finding a subgraph isomorphism $i$ from a graph $G$ to a graph
    $H$, the following constraints are implied for any pair of vertices $v, w \in \vertexset(G)$:
    \begin{enumerate}
        \item The distance between $v$ and $w$ is at most the distance between $i(v)$ and $i(w)$.
        \item For any pair of vertices $v, w \in \vertexset(G)$, if there are at least $k$ simple
            paths of length exactly $\ell$ between $v$ and $w$, then there must be at least $k$
            simple paths of length exactly $\ell$ between $i(v)$ and $i(w)$.
    \end{enumerate}
\end{proposition}

?? This fourth one is hard. The third one actually works for homomorphisms.

?? Corollary: $G^d$ graphs.

?? Clique, be very careful with target loops. What we want to prove is, if the maximum clique in the
neighbourhood of a pattern vertex $p$ has $n$ vertices, and if the target graph has no loops, then
$p$ cannot be mapped to any target vertex $t$ unless $t$ is contained in a clique of at least $n$
vertices.

?? The point is we can solve clique decision much faster than general homomorphism, and we only need
to do it once per target vertex.

\section{Design and Implementation}

We use the subgraph isomorphism code in the Glasgow Subgraph Solver \cite{DBLP:conf/gg/McCreeshP020}
as a starting point.  We implemented distance filtering for the basic homomorphism problem, using
the existing supplemental graphs framework that the solver uses for subgraph isomorphism to create
pseudo-adjacency constraints for vertices that are distance two away from each other; this was
routine.  We also implemented clique constraints, initially using the Glasgow Subgraph Solver's
existing maximum clique solver to perform domain filtering. Preliminary experiments showed that a
na{\"\i}ve approach, which begins by calculating the maximum clique size for the neighbourhood of
each pattern vertex and each target vertex, would add as much as three minutes of preprocessing time
to some problem instances which could otherwise be solved in a few seconds. We therefore use a more
intelligent approach, which we now describe.

\subsection{Optimising Maximum Clique Solver Calls}

We calculate the maximum clique size for the neighbourhood of a first pattern vertex $p_1$.  Having
found a maximum clique $\omega_1$ for a vertex $p$, for every other vertex $q \in \omega_1$, we
remember that the maximum clique for $q$'s neighbourhood has at least $|\omega_1|$ vertices. We then
proceed to find a maximum clique $\omega_2$ for a second pattern vertex $p_2$; however, if $p_2$ was
a member of $\omega_1$, we initialise the clique solver with a lower bound of $n$ rather than
starting from scratch. We then find the maximum clique size for a third pattern vertex $p_3$, using
whichever lower bound of $|\omega_1|$ and $|\omega_2|$ is better (if both are valid), and so on.

We then move on to the target vertices, using a similar caching routine. Rather than calculating a
clique size for every single target vertex, we only calculate a value for target vertices which are
present in at least one variable's domain, after other unary constraints have been applied.
(Depending upon the problem variant, implied constraints like degree and neighbourhood degree
sequence can eliminate some target vertices from the domains of all pattern variables.)
Additionally, we do not require the maximum clique solver to run to completion and guarantee that it
has found a maximum clique. Instead, we allow it to stop as soon as it has found a clique with as
many vertices as the largest pattern clique. This is useful in practice because it may be very hard
to decide whether a particular target vertex has neighbourhood clique size of, say, 15 or 16, but if
the largest pattern vertex has a neighbourhood clique size of only 5 then this is irrelevant.

?? Why this works

\subsection{Proof Logging}

?? How to log proofs

\section{Experiments}

\begin{figure*}[p]
    \,\hfill\begin{tabular}{c@{}c@{}c@{}}
    \includegraphics{gen-graph-cumulatives.pdf}
        &
    \includegraphics{gen-graph-cumulatives-differences.pdf}
        &
    \includegraphics{gen-graph-cumulatives-aggregate.pdf}
        \\[5mm]
    \includegraphics{gen-graph-cumulatives-hom.pdf}
        &
    \includegraphics{gen-graph-cumulatives-local.pdf}
        &
    \includegraphics{gen-graph-cumulatives-si.pdf}
    \end{tabular}\hfill\,

    \caption{On the top left, the cumulative number of instances solved over time for the three
    problem variants, with and without clique filtering, and with and without distance filtering for
    the homomorphism problem. The remaining plots re-display this data, as follows. The three plots
    on the bottom row zoom in on the cumulative number of instances solved, for the homomorphism
    problem on the left, the locally injective homomorphism problem in the centre, and the subgraph
    isomorphism problem on the right. The top centre plot shows the additional number of instances
    solved at any given time when using the new forms of filtering for all problem variants, and the
    top right plot shows the aggregate speedups from each form of filtering.}
\end{figure*}
\begin{figure*}[p]
    \,\hfill\begin{tabular}{c@{\hspace{4mm}}c@{\hspace{4mm}}c@{\hspace{4mm}}c@{\hspace{4mm}}}
        \includegraphics{gen-graph-scatter-hom-c.pdf}
        &
        \includegraphics{gen-graph-scatter-hom-cd.pdf}
        &
        \includegraphics{gen-graph-scatter-local.pdf}
        &
        \includegraphics{gen-graph-scatter-si.pdf}
    \end{tabular}\hfill\,

    \caption{Looking at the effects of additional filtering on an instance by instance basis, for
    homomorphism with just clique filtering and with both clique and distance filtering, and for the
    other two variants with clique filtering. Each point represents one instance, the vertical axis
    is the runtime with filtering in ms, and the horizontal axis is the runtime without filtering in
    ms (and so points below the diagonal are speedups). Points on the outer axes are timeouts. The
    different point styles show the different families of instance from the benchmark set, and
    illustrate that in each the filtering is broadly useful rather than being specific to a single
    kind of application.}
\end{figure*}

We now evaluate this approach empirically.

?? How good is it

?? Time on cliques

?? Whether cliques are hard

\section{Conclusion and Future Work}

?? Neighbourhood colourability? Cores?

\section*{Acknowledgements}

This work was supported by the Engineering and Physical Sciences Research Council [grant number
EP/P026842/1].

\bibliographystyle{named}
\bibliography{paper}

\end{document}

% vim: set tw=100 spell spelllang=en : %

### 2021-01-09 paper/aggregates.cc

@@ -1,0 +1,93 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include <cstdlib>
+#include <fstream>
+#include <iostream>
+#include <optional>
+#include <string>
+#include <utility>
+#include <vector>
+
+using std::cout;
+using std::endl;
+using std::ifstream;
+using std::nullopt;
+using std::optional;
+using std::pair;
+using std::string;
+using std::vector;
+
+auto get(ifstream & s) -> optional<pair<unsigned, unsigned> >
+{
+    string tok, tok2, _;
+    while (true) {
+        if (! (s >> tok))
+            return nullopt;
+        if (tok != "#")
+            break;
+        getline(s, _);
+    }
+    s >> tok2;
+    getline(s, _);
+    return pair{ stoi(tok), stoi(tok2) };
+}
+
+auto main(int argc, char * argv[]) -> int
+{
+    if (3 != argc)
+        return EXIT_FAILURE;
+
+    ifstream f1{ argv[1] }, f2{ argv[2] };
+
+    vector<unsigned> v1, v2;
+    for (auto d1 = get(f1) ; d1 ; d1 = get(f1)) {
+        while (v1.size() < d1->first)
+            v1.push_back(v1.back());
+        v1.push_back(d1->second);
+    }
+
+    for (auto d2 = get(f2) ; d2 ; d2 = get(f2)) {
+        while (v2.size() < d2->first)
+            v2.push_back(v2.back());
+        v2.push_back(d2->second);
+    }
+
+    while (v1.size() < v2.size())
+        v1.push_back(v1.back());
+    while (v2.size() < v1.size())
+        v2.push_back(v2.back());
+
+    unsigned n1 = 0, t1 = 0, n2 = 0, t2 = 0, t12 = 0, t21 = 0;
+    for (unsigned i = 0 ; i < v1.size() ; ++i) {
+        if (v1[i] > n1) {
+            n1 = v1[i];
+            t1 = i;
+        }
+
+        if (v2[i] > n2) {
+            n2 = v2[i];
+            t2 = i;
+        }
+
+        if (n1 > n2) {
+            // how long did 1 take to solve n2 instances?
+            while (v1[t12] < n2)
+                ++t12;
+            if (t2 > 1 && t12 > 1)
+                cout << i << " " << (double(t12) / double(t2)) << endl;
+        }
+        else if (n1 < n2) {
+            // how long did 2 take to solve n1 instances?
[…]

### 2021-01-09 paper/differences.cc

@@ -1,0 +1,65 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include <cstdlib>
+#include <fstream>
+#include <iostream>
+#include <optional>
+#include <string>
+#include <utility>
+#include <vector>
+
+using std::cout;
+using std::endl;
+using std::ifstream;
+using std::nullopt;
+using std::optional;
+using std::pair;
+using std::string;
+using std::vector;
+
+auto get(ifstream & s) -> optional<pair<unsigned, unsigned> >
+{
+    string tok, tok2, _;
+    while (true) {
+        if (! (s >> tok))
+            return nullopt;
+        if (tok != "#")
+            break;
+        getline(s, _);
+    }
+    s >> tok2;
+    getline(s, _);
+    return pair{ stoi(tok), stoi(tok2) };
+}
+
+auto main(int argc, char * argv[]) -> int
+{
+    if (3 != argc)
+        return EXIT_FAILURE;
+
+    ifstream f1{ argv[1] }, f2{ argv[2] };
+
+    vector<unsigned> v1, v2;
+    for (auto d1 = get(f1) ; d1 ; d1 = get(f1)) {
+        while (v1.size() < d1->first)
+            v1.push_back(v1.back());
+        v1.push_back(d1->second);
+    }
+
+    for (auto d2 = get(f2) ; d2 ; d2 = get(f2)) {
+        while (v2.size() < d2->first)
+            v2.push_back(v2.back());
+        v2.push_back(d2->second);
+    }
+
+    while (v1.size() < v2.size())
+        v1.push_back(v1.back());
+    while (v2.size() < v1.size())
+        v2.push_back(v2.back());
+
+    for (unsigned i = 0 ; i < v1.size() ; ++i)
+        cout << i << " " << v1[i] << " " << v2[i] << endl;
+
+    return EXIT_SUCCESS;
+}
+

### 2021-01-11 paper/paper.tex
More writing, yay

@@
 
+\usepackage{cleveref}
+
 % \usepackage{showframe}
@@
 \newtheorem{proposition}{Proposition}
+\newtheorem{corollary}{Corollary}
 
@@
 
-\title{Solving Graph Homomorphism and Subgraph Isomorphism Problems Faster Through Neighbourhood
-Maximum Clique Constraints}
+\title{Solving Graph Homomorphism and Subgraph Isomorphism Problems Faster Through Clique Neighbourhood
+Constraints}
 
@@
 
-Let $G$ and $H$ be graphs. Let $v \in \vertexset(G)$ be a vertex of $G$. The \emph{(open)
-neighbourhood} of $v$, written $\neighbourhood_G(v)$, is the set of vertices adjacent to $v$ not
-including $v$ itself, whilst the closed neighbourhood of $v$, written $\neighbourhood_G[v]$, is the
-neighbourhood plus $v$. The \emph{degree} of a vertex, $\degree_G(v)$, is the cardinality of its open
-neighbourhood. Given a vertex set $S \subseteq \vertexset(G)$, the subgraph \emph{induced by} $S$,
-written $G[S]$, is the subgraph of $G$ with only the vertices in $S$ together with all the edges
-between them. A \emph{clique} is a subgraph where every vertex is adjacent to every other in the
-subgraph.
+We begin by introducing notation and terminology, and by giving the theoretical results which
+support our implementation.
 
-A \emph{homomorphism} from $G$ to $H$ is a function mapping vertices of $G$ to vertices of $H$, such
-that adjacent vertices in $G$ are mapped to adjacent vertices in $H$. A homomorphism $h$ is
-\emph{locally injective} if for every vertex $w$, the restriction of $h$ to $G[N_G[w]]$ is
-injective; if $h$ is injective globally we call it a \emph{(non-induced) subgraph isomorphism}.
+\paragraph{Graphs.} Let $G$ and $H$ be graphs. Let $v \in \vertexset(G)$ be a vertex of $G$. The
+\emph{(open) neighbourhood} of $v$, written $\neighbourhood_G(v)$, is the set of vertices adjacent
+to $v$ not including $v$ itself, whilst the closed neighbourhood of $v$, written
+$\neighbourhood_G[v]$, is the neighbourhood plus $v$. The \emph{degree} of a vertex, $\degree_G(v)$,
+is the cardinality of its open neighbourhood. The \emph{neighbourhood degree sequence} of a vertex
+is the sequence consisting of the degrees of its neighbours, in descending order. Given a vertex set
+$S \subseteq \vertexset(G)$, the subgraph \emph{induced by} $S$, written $G[S]$, is the subgraph of
+$G$ with only the vertices in $S$ together with all the edges between them. A \emph{clique} is a
+subgraph where every vertex is adjacent to every other in the subgraph.
+
+\paragraph{Homomorphisms.} A \emph{homomorphism} from $G$ to $H$ is a function mapping vertices of
+$G$ to vertices of $H$, such that adjacent vertices in $G$ are mapped to adjacent vertices in $H$. A
+homomorphism $h$ is \emph{locally injective} if for every vertex $w$, the restriction of $h$ to
+$G[N_G[w]]$ is injective; if $h$ is injective globally we call it a \emph{(non-induced) subgraph
+isomorphism}.
 
@@
 
-?? CSP, hom as CSP
+\paragraph{Constraint programming.} ?? A constraint programming algorithm.
 
-An \emph{implied constraint} is a constraint that can be added to a CSP without altering its set of
-solutions. ?? Is it best wording all of this in terms of implied constraints?
+\paragraph{Degree-based invariants for homomorphisms.} The use of constraint programming techniques
+for finding homomorphisms relies upon exploiting various invariants. For example, every recent
+constraint programming solver makes use of the following result.
 
-\begin{proposition}For the problem of finding a subgraph isomorphism $i$ from a graph $G$ to a graph
-    $H$, the following constraints are implied for any vertex $v \in \vertexset(G)$:
+\begin{proposition}\label{proposition:degreends}
+    Suppose a subgraph isomorphism $i$ exists from a graph $G$ to a graph $H$. Let
+    $v$ be any vertex. Then:
     \begin{enumerate}
-        \item We cannot map $v$ to a vertex of lower degree, $\degree(v) \le \degree(i(v))$.
-        \item Furthermore, ??NDS.
+        \item The function cannot map $v$ to a vertex of lower degree, $\degree(v) \le \degree(i(v))$.
+        \item Furthermore, $v$ cannot be mapped to any vertex whose neighbourhood degree sequence is
+            not pointwise greater than or equal to its own
+            \cite{DBLP:journals/constraints/ZampelliDS10}.
     \end{enumerate}
@@
 
-?? Show that these work for locally injective but not homomorphism
+The simplest way for a constraint programming algorithm to use these results is as unary
+constraints, which are propagated when domains are initialised, before search begins. For each
+domain representing a pattern vertex, any value representing target vertex whose degree or
+neighbourhood degree sequence is too low may immediately be rejected.
+
+These invariants do \emph{not} hold for homomorphisms in general. As a simple counter-example, a
+homomorphism may map a star graph (one central vertex of high degree, which is adjacent to many
+other vertices, each of which has no other neighbours) onto a single edge. However, it is
+straightforward to verify the following.
+
+\begin{proposition}Both invariants in \cref{proposition:degreends} are also valid for locally
+injective homomorphisms.\end{proposition}
+
+\paragraph{Path-based invariants for homomorphisms.} Constraint programming solvers can also exploit
+invariants that are based upon paths. This is done by automatically adding \emph{implied}
+constraints to the problem that are implied by the original model, but that will give stronger
+propagation power.
 
@@
     \begin{enumerate}
-        \item The distance between $v$ and $w$ is at most the distance between $i(v)$ and $i(w)$.
-        \item For any pair of vertices $v, w \in \vertexset(G)$, if there are at least $k$ simple
+        \item The distance between $v$ and $w$ is at most the distance between $i(v)$ and $i(w)$
+            \cite{DBLP:conf/cp/AudemardLMGP14}.
+        \item If there are at least $k$ simple
             paths of length exactly $\ell$ between $v$ and $w$, then there must be at least $k$
-            simple paths of length exactly $\ell$ between $i(v)$ and $i(w)$.
+            simple paths of length exactly $\ell$ between $i(v)$ and $i(w)$
+            \cite{DBLP:conf/cp/McCreeshP15}.
     \end{enumerate}
@@
 
-?? This fourth one is hard. The third one actually works for homomorphisms.
+It is easy to verify that the first of these two properties is valid for \emph{any} homomorphism.
+The second property does not hold for homomorphisms in general: for example, a pair of vertices
+connected by three paths of length two may be mapped onto a pair of vertices connected by a single
+path of length two. Local injectivity is trickier. The second property does not hold in full
+generality, because, for example, a path of length three may be mapped onto a triangle whilst
+respecting local injectivity. However, a weaker result holds:
+
+\begin{proposition}For the problem of finding a locally injective graph homomorphism $i$ from a
+    graph $G$ to a graph $H$, for any pair of vertices $v, w \in \vertexset(G)$, if there are at
+    least $k$ simple paths of length exactly two between $v$ and $w$, then there must be at least
+    $k$ simple paths of length exactly two between $i(v)$ and $i(w)$.
+\end{proposition}
+
+\begin{proof}Let $\{ x_1, \ldots, x_n \}$ be the intermediate vertices on the paths of length two
+between $v$ and $w$. Observe that each $x_j$ is in the neighbourhood of $v$, and so must be mapped
+    to different vertices. Thus each sequence $(i(v), i(x_i), i(w))$ gives a distinct simple path of
+length two between $i(v)$ and $i(w)$ as required.\end{proof}
 
@@
 
-?? Clique, be very careful with target loops. What we want to prove is, if the maximum clique in the
-neighbourhood of a pattern vertex $p$ has $n$ vertices, and if the target graph has no loops, then
-$p$ cannot be mapped to any target vertex $t$ unless $t$ is contained in a clique of at least $n$
-vertices.
+\paragraph{Cliques and homomorphisms.} Finally, we come to the main result of this paper. Observe
+that until now, we have not seen any unary constraints which are valid for the general homomorphism
+problem. All existing constraint programming approaches for the subgraph isomorphism problem begin
+by branching on the variable which has the smallest domain---typically this will be a domain for a
+pattern vertex of high degree, or which has many high degree neighbours. Furthermore, once one such
+guessed assignment has been made, adjacency propagation means many other domains will be
+substantially reduced in size, making subsequent branching choices simpler.  However, for
+homomorphism problems, if we cannot find any implied unary constraints then every domain will
+initially be of the same size, which will make it much harder for a solver to know where to start.
+This motivates the following observation.
 
-?? The point is we can solve clique decision much faster than general homomorphism, and we only need
-to do it once per target vertex.
+\begin{proposition}\label{proposition:clique}
+    Suppose we are finding a graph homomorphism $i$ from a graph $G$ to a graph $H$ where $H$ does
+    not contain any loops. Suppose there is a $k$ vertex clique in $G$. Then the image of this
+    clique under $i$ must also be a $k$ vertex clique.
+\end{proposition}
+
+\begin{proof}
+    Let $j$ and $k$ be two distinct vertices of the clique in $G$. By definition of a homomorphism,
+    $i(j)$ and $i(k)$ must be adjacent. And, because $H$ does not contain loops, $i(j)$ cannot equal
+    $i(k)$.
+\end{proof}
+
+\begin{corollary}\Cref{proposition:clique} holds for locally injective graph homomorphisms and for
+    subgraph isomorphisms even if loops are present in the second graph.
+\end{corollary}
+
+We will discuss how this is useful in practice in the following section.
 
@@
 
-We use the subgraph isomorphism code in the Glasgow Subgraph Solver \cite{DBLP:conf/gg/McCreeshP020}
-as a starting point.  We implemented distance filtering for the basic homomorphism problem, using
+Having determined which commonly-used subgraph isomorphism invariants do and do not hold for other
+forms of homomorphism, and having discovered a new clique-based invariant, we will now look at how
+these invariants may be used in practice. Our starting point is the Glasgow Subgraph Solver
+\cite{DBLP:conf/gg/McCreeshP020}, due to it being the current single strongest solver for the
+subgraph isomorphism problem.
+
+??
+
+We implemented distance filtering for the basic homomorphism problem, using
 the existing supplemental graphs framework that the solver uses for subgraph isomorphism to create
 pseudo-adjacency constraints for vertices that are distance two away from each other; this was
-routine.  We also implemented clique constraints, initially using the Glasgow Subgraph Solver's
-existing maximum clique solver to perform domain filtering. Preliminary experiments showed that a
-na{\"\i}ve approach, which begins by calculating the maximum clique size for the neighbourhood of
-each pattern vertex and each target vertex, would add as much as three minutes of preprocessing time
-to some problem instances which could otherwise be solved in a few seconds. We therefore use a more
+routine.
+
+We also implemented clique constraints, initially using the Glasgow Subgraph Solver's existing
+maximum clique solver to perform domain filtering. Preliminary experiments showed that a na{\"\i}ve
+approach, which begins by calculating the maximum clique size for the neighbourhood of each pattern
+vertex and each target vertex, would add as much as three minutes of preprocessing time to some
+problem instances which could otherwise be solved in a few seconds. We therefore use a more
 intelligent approach, which we now describe.

### 2021-01-12 paper/paper.tex
Changes from Sonja in Overleaf

@@
 to $v$ not including $v$ itself, whilst the closed neighbourhood of $v$, written
-$\neighbourhood_G[v]$, is the neighbourhood plus $v$. The \emph{degree} of a vertex, $\degree_G(v)$,
+$\neighbourhood_G[v]$, is the neighbourhood of $v$ plus $v$. The \emph{degree} of a vertex, $\degree_G(v)$,
 is the cardinality of its open neighbourhood. The \emph{neighbourhood degree sequence} of a vertex
@@
 \paragraph{Degree-based invariants for homomorphisms.} The use of constraint programming techniques
-for finding homomorphisms relies upon exploiting various invariants. For example, every recent
-constraint programming solver makes use of the following result.
+for finding homomorphisms relies upon exploiting various invariants. It is straightforward to verify the following.
 
 \begin{proposition}\label{proposition:degreends}
-    Suppose a subgraph isomorphism $i$ exists from a graph $G$ to a graph $H$. Let
+    Suppose  $i$ is a locally injective homomorphism from a $G$ to  $H$. Let
     $v$ be any vertex. Then:
     \begin{enumerate}
-        \item The function cannot map $v$ to a vertex of lower degree, $\degree(v) \le \degree(i(v))$.
+        \item $i$ cannot map $v$ to a vertex of lower degree, $\degree_G(v) \le \degree_H(i(v))$.
         \item Furthermore, $v$ cannot be mapped to any vertex whose neighbourhood degree sequence is
             not pointwise greater than or equal to its own
-            \cite{DBLP:journals/constraints/ZampelliDS10}.
+           .
     \end{enumerate}
 \end{proposition}
+Since subgraph isomorphisms are in particular locally injective, these also hold for subgraph isomorphisms \cite{DBLP:journals/constraints/ZampelliDS10}.
+Every recent
+constraint programming solver makes use of the of the above invariants for subgraph isomowphisms.
+However, these invariants do \emph{not} hold for homomorphisms in general. As a simple counter-example, a
+homomorphism may map a star graph (one central vertex of high degree, which is adjacent to many
+other vertices, each of which has no other neighbours) onto a single edge.
 
@@
 neighbourhood degree sequence is too low may immediately be rejected.
-
-These invariants do \emph{not} hold for homomorphisms in general. As a simple counter-example, a
-homomorphism may map a star graph (one central vertex of high degree, which is adjacent to many
-other vertices, each of which has no other neighbours) onto a single edge. However, it is
-straightforward to verify the following.
-
-\begin{proposition}Both invariants in \cref{proposition:degreends} are also valid for locally
-injective homomorphisms.\end{proposition}
 
@@
     \begin{enumerate}
-        \item The distance between $v$ and $w$ is at most the distance between $i(v)$ and $i(w)$
+        \item The distance between $v$ and $w$ is at least the distance between $i(v)$ and $i(w)$
             \cite{DBLP:conf/cp/AudemardLMGP14}.
@@
 \begin{proposition}\label{proposition:clique}
-    Suppose we are finding a graph homomorphism $i$ from a graph $G$ to a graph $H$ where $H$ does
-    not contain any loops. Suppose there is a $k$ vertex clique in $G$. Then the image of this
-    clique under $i$ must also be a $k$ vertex clique.
+    Let $i$ be a homomorphism from $G$ to $H$ where $H$ does
+    not contain any loops. Let $S\subset V$ be such that $G[S]=K_k$
 \end{proposition}

### 2021-01-13 paper/paper.tex
More from Sonja
Better words

@@
 
-\title{Solving Graph Homomorphism and Subgraph Isomorphism Problems Faster Through Clique Neighbourhood
-Constraints}
+\title{Solving Graph Homomorphism and Subgraph Isomorphism Problems Faster Through Clique Neighbourhood Constraints}
 
@@
 \begin{abstract}
-    We can find graph homomorphisms faster through clique constraints.
+    Graph homomorphism problems involve finding adjacency-preserving mappings between two graphs.
+    Many of these problems are theoretically hard, but can be solved quite efficiently in practice
+    through constraint programming techniques.  This paper adapts techniques from the
+    state-of-the-art in subgraph isomorphism solving to broader graph homomorphism problems, and
+    introduces a new form of filtering based upon clique-finding. We evaluate these techniques
+    empirically, and show they are effective in practice for the locally injective graph
+    homomorphism and subgraph isomorphism problems, as well as giving the first practical constraint
+    programming approach to finding general graph homomorphisms.
 \end{abstract}
@@
 variants. Indeed, this paper proves that many, but not all, of these strategies are also valid in
-the locally injective case, and that a weakened form of one other strategy is valid for all
-homomorphisms. However, we also prove that \emph{none} of the other strategies are valid for finding
+the locally injective case, and that simple distance filtering is valid for all
+homomorphisms. However, we also show that \emph{none} of the other strategies are valid for finding
 homomorphisms where there is no injectivity requirement. Finally, we introduce a new filtering
-technique that is based upon finding the maximum clique in the neighbourhood of each vertex, and
+technique that is based upon finding a maximum clique in the neighbourhood of each pattern vertex, and
 prove that this technique is valid even in the general case. Although this new filtering technique
 involves solving a large number of \NP-complete problems, we demonstrate that it is effective in
-practice, particularly for the non-injective problem where filtering can solve instances ??many
-hundreds of times faster in aggregate. This shows, for the first time, that modern constraint
-programming techniques can be practical for less constrained graph homomorphism finding problems;
-previous algorithmic approaches have focused instead upon worst-case computational bounds
+practice, particularly for the non-injective problem where filtering allows a constraint programming
+algorithm to solve instances over eight hundred times faster in aggregate, over a collection of over
+fourteen thousand standard benchmark instances. This shows, for the first time, that modern
+constraint programming techniques can be practical for less constrained graph homomorphism finding
+problems; previous algorithmic approaches have focused instead upon worst-case computational bounds
 \cite{DBLP:journals/csr/FialaK08,DBLP:journals/ipl/Rzazewski14,DBLP:journals/tcs/ChaplickFHPT15},
@@
 Every recent
-constraint programming solver makes use of the of the above invariants for subgraph isomowphisms.
+constraint programming solver makes use of the of the above invariants for subgraph isomorphisms.
 However, these invariants do \emph{not} hold for homomorphisms in general. As a simple counter-example, a
@@
 
-\begin{proposition}\label{proposition:clique}
+\begin{proposition}[Cliques are preserved]\label{proposition:clique}
     Let $i$ be a homomorphism from $G$ to $H$ where $H$ does
-    not contain any loops. Let $S\subset V$ be such that $G[S]=K_k$
+    not contain any loops. Let $S\subset V$ be such that $G[S]=K_k$. Then $H[i(S)]=K_k$. 
 \end{proposition}

### 2021-01-15 paper/paper.tex
More writings

@@
 \usepackage{cleveref}
+\crefname{figure}{figure}{figures}
+\Crefname{figure}{Figure}{Figures}
+
+\usepackage{tikz}
+\usetikzlibrary{shapes}
 
@@
 \newtheorem{corollary}{Corollary}
+
+\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
 
@@
 
-?? loops, be really careful with this
+\paragraph{Loops.} A vertex which is adjacent to itself is called a \emph{loop}. By a careful reading
+of the definition, any homomorphism must map loops onto loops---and indeed, the Glasgow Subgraph
+Solver and this paper take this approach (although some other algorithm implementations do not). It
+therefore follows that for the homomorphism \emph{decision} problem where there are no injectivity
+constraints, any problem instance which has a loop in the codomain graph is trivially satisfiable,
+although the \emph{counting} problem remains \#\P-hard \cite{DBLP:journals/rsa/DyerG00}.
 
-\paragraph{Constraint programming.} ?? A constraint programming algorithm.
+\paragraph{Constraint programming.} A \emph{constraint satisfaction problem} is defined in terms of
+a set of \emph{variables}, each of which has a \emph{domain} of possible \emph{values}, together
+with a set of constraints; the goal is to give each variable a value from its domain, whilst
+satisfying all of the constraints. Homomorphism problems have a natural representation as a
+constraint satisfaction problem: we have a variable for each vertex in the domain graph, whose
+values range over the codomain graph, and a set of constraints saying that adjacent vertices must be
+mapped to adjacent vertices. For the injective variants, one or more \emph{all-different}
+constraints are also present. A typical constraint programming approach to solving such a problem is
+to combine inference through constraint propagation with an intelligent backtracking search.
 
-\paragraph{Degree-based invariants for homomorphisms.} The use of constraint programming techniques
-for finding homomorphisms relies upon exploiting various invariants. It is straightforward to verify the following.
+\paragraph{Degree-based invariants for homomorphisms.} Constraint programming techniques for finding
+homomorphisms can be made much more effective by exploiting certain invariants. It is
+straightforward to verify the following.
 
@@
         \item Furthermore, $v$ cannot be mapped to any vertex whose neighbourhood degree sequence is
-            not pointwise greater than or equal to its own
-           .
+            not pointwise greater than or equal to its own.
     \end{enumerate}
 \end{proposition}
-Since subgraph isomorphisms are in particular locally injective, these also hold for subgraph isomorphisms \cite{DBLP:journals/constraints/ZampelliDS10}.
-Every recent
-constraint programming solver makes use of the of the above invariants for subgraph isomorphisms.
-However, these invariants do \emph{not} hold for homomorphisms in general. As a simple counter-example, a
-homomorphism may map a star graph (one central vertex of high degree, which is adjacent to many
-other vertices, each of which has no other neighbours) onto a single edge.
-
-The simplest way for a constraint programming algorithm to use these results is as unary
+Since subgraph isomorphisms are in particular locally injective, these invariants also hold for
+subgraph isomorphisms \cite{DBLP:journals/constraints/ZampelliDS10}; indeed, every recent constraint
+programming approach for subgraph isomorphism finding make use of degrees and neighbourhood degree
+sequences. The simplest way for a constraint programming algorithm to use these results is as unary
 constraints, which are propagated when domains are initialised, before search begins. For each
@@
 neighbourhood degree sequence is too low may immediately be rejected.
+
+\begin{figure}
+    \hfill\,\begin{minipage}{3cm}
+    \begin{tikzpicture}
+        \node [draw, fill, circle, inner sep=1.5pt] (A1) at (0, 0) {};
+        \node [draw, circle, inner sep=1.5pt] (A2) at (0:0.3) {};
+        \node [draw, circle, inner sep=1.5pt] (A3) at (60:0.3) {};
+        \node [draw, circle, inner sep=1.5pt] (A4) at (120:0.3) {};
+        \node [draw, circle, inner sep=1.5pt] (A5) at (180:0.3) {};
+        \node [draw, circle, inner sep=1.5pt] (A6) at (240:0.3) {};
+        \node [draw, circle, inner sep=1.5pt] (A7) at (300:0.3) {};
+
+        \draw (A1) -- (A2);
+        \draw (A1) -- (A3);
+        \draw (A1) -- (A4);
+        \draw (A1) -- (A5);
+        \draw (A1) -- (A6);
+        \draw (A1) -- (A7);
+
+        \draw [->] (0.7, 0) -- (1.1, 0);
+
+        \draw (1.5, 0) node [draw, fill, circle, inner sep=1.5pt] (B1) at ++ (90:0.15) {};
+        \draw (1.5, 0) node [draw, circle, inner sep=1.5pt] (B2) at ++ (270:0.15) {};
+        \draw (B1) -- (B2);
+    \end{tikzpicture}\end{minipage}\hfill\begin{minipage}{3cm}
+    \begin{tikzpicture}
+        \node [draw, fill, circle, inner sep=1.5pt] (A1) at (0, 0) {};
+        \node [draw, inner sep=1.5pt] (A2) at (0.3, 0) {};
+        \node [draw, inner sep=1.5pt] (A3) at (0.3, -0.2) {};
+        \node [draw, inner sep=1.5pt] (A4) at (0.3, 0.2) {};
+        \node [draw, circle, inner sep=1.5pt] (A5) at (0.6, 0) {};
+
+        \draw (A1) -- (A2);
+        \draw (A1) -- (A3);
+        \draw (A1) -- (A4);
+        \draw (A2) -- (A5);
+        \draw (A3) -- (A5);
+        \draw (A4) -- (A5);
+
+        \draw [->] (1.0, 0) -- (1.4, 0);
+
+        \node [draw, fill, circle, inner sep=1.5pt] (B1) at (1.8, 0) {};
+        \node [draw, inner sep=1.5pt] (B2) at (2.1, 0) {};
+        \node [draw, circle, inner sep=1.5pt] (B5) at (2.4, 0) {};
+
+        \draw (B1) -- (B2);
+        \draw (B2) -- (B5);
+    \end{tikzpicture}\end{minipage}\hfill\,
+
+    \bigskip
+
+    \,\hfill\begin{minipage}{3cm}
+    \begin{tikzpicture}
+        \node [draw, fill, circle, inner sep=1.5pt] (A1) at (0, 0) {};
+        \node [draw, circle, inner sep=1.5pt] (A2) at (0.3, 0) {};
+        \node [draw, inner sep=1.5pt] (A3) at (0.6, 0) {};
+        \node [draw, fill, circle, inner sep=1.5pt] (A4) at (0.9, 0) {};
+
+        \draw (A1) -- (A2);
+        \draw (A2) -- (A3);
+        \draw (A3) -- (A4);
+
+        \draw [->] (1.3, 0) -- (1.7, 0);
+
+        \draw (2.2, 0) node [draw, fill, circle, inner sep=1.5pt] (B1) at ++ (120:0.3) {};
+        \draw (2.2, 0) node [draw, circle, inner sep=1.5pt] (B2) at ++ (240:0.3) {};
+        \draw (2.2, 0) node [draw, inner sep=1.5pt] (B3) at ++ (0:0.3) {};
+
+        \draw (B1) -- (B2);
+        \draw (B2) -- (B3);
+        \draw (B3) -- (B1);
+    \end{tikzpicture}\end{minipage}\hfill\begin{minipage}{3cm}
+        \begin{tikzpicture}
+        \node [draw, fill, circle, inner sep=1.5pt] (A1) at (0, 0) {};
+        \node [draw, inner sep=1.5pt] (A2) at (0.3, -0.2) {};
+        \node [draw, fill, inner sep=1.5pt] (A3) at (0.3, 0.2) {};
+        \node [draw, circle, inner sep=1.5pt] (A4) at (0.6, 0) {};
+        \node [draw, diamond, fill, inner sep=1.5pt] (A5) at (0.9, -0.2) {};
+        \node [draw, diamond, inner sep=1.5pt] (A6) at (0.9, 0.2) {};
+        \node [draw, fill, circle, inner sep=1.5pt] (A7) at (1.2, 0) {};
+
+        \draw (A1) -- (A2);
+        \draw (A1) -- (A3);
+        \draw (A2) -- (A4);
+        \draw (A3) -- (A4);
+        \draw (A4) -- (A5);
+        \draw (A4) -- (A6);
+        \draw (A5) -- (A7);
+        \draw (A6) -- (A7);
+
+        \draw [->] (1.4, 0) -- (1.7, 0);
+
+        \node [draw, fill, circle, inner sep=1.5pt] (B1) at (2, 0) {};
+        \node [draw, fill, inner sep=1.5pt] (B2) at (2.3, -0.15) {};
+        \node [draw, diamond, inner sep=1.5pt] (B3) at (2.3, 0.15) {};
+        \node [draw, diamond, fill, inner sep=1.5pt] (B4) at (2.3, -0.45) {};
+        \node [draw, inner sep=1.5pt] (B5) at (2.3, 0.45) {};
+        \node [draw, circle, inner sep=1.5pt] (B6) at (2.6, 0) {};
+
+        \draw (B1) -- (B2);
+        \draw (B1) -- (B3);
+        \draw (B1) -- (B4);
+        \draw (B1) -- (B5);
+        \draw (B6) -- (B2);
+        \draw (B6) -- (B3);
+        \draw (B6) -- (B4);
+        \draw (B6) -- (B5);
+    \end{tikzpicture}\end{minipage}\hfill\,
+
+    \caption{Counter-examples for various properties which are not invariants. The different shapes
+        for vertices show a mapping, with vertices being mapped to vertices of the same shape. The
+        top left example shows that degree is not preserved in a homomorphism; the top right that path
+        counts are not preserved in a homomorphism; the bottom left that path counts of length three
+        are not preserved in a locally injective homomorphism; and the bottom right that a locally
+        injective homomorphism $i$ can give rise to a homomorphism $i^{2,2}$ which is not locally
+        injective.} \label{figure:counterexamples}
+\end{figure}
+
+However, these invariants do \emph{not} hold for homomorphisms in general. As a simple
+counter-example, a homomorphism may map a star graph onto a single edge, as in
+\cref{figure:counterexamples} (top left).
 
@@
 
-\begin{proposition}For the problem of finding a subgraph isomorphism $i$ from a graph $G$ to a graph
-    $H$, the following constraints are implied for any pair of vertices $v, w \in \vertexset(G)$:
+\begin{proposition}\label{proposition:distances}For the problem of finding a subgraph isomorphism
+    $i$ from a graph $G$ to a graph $H$, the following constraints are implied for any pair of
+    vertices $v, w \in \vertexset(G)$:
     \begin{enumerate}
@@
 connected by three paths of length two may be mapped onto a pair of vertices connected by a single
-path of length two. Local injectivity is trickier. The second property does not hold in full
-generality, because, for example, a path of length three may be mapped onto a triangle whilst
-respecting local injectivity. However, a weaker result holds:
+path of length two (see \cref{figure:counterexamples}, top right). Local injectivity is trickier.
+The second property does not hold in full generality, because, for example, a path of length three
+may be mapped onto a triangle whilst respecting local injectivity (\cref{figure:counterexamples},
+bottom left).  However, we observe that a weaker result holds:
 
-\begin{proposition}For the problem of finding a locally injective graph homomorphism $i$ from a
-    graph $G$ to a graph $H$, for any pair of vertices $v, w \in \vertexset(G)$, if there are at
-    least $k$ simple paths of length exactly two between $v$ and $w$, then there must be at least
-    $k$ simple paths of length exactly two between $i(v)$ and $i(w)$.
+\begin{proposition}\label{proposition:paths}For the problem of finding a locally injective graph
+    homomorphism $i$ from a graph $G$ to a graph $H$, for any pair of vertices $v, w \in
+    \vertexset(G)$, if there are at least $k$ simple paths of length exactly two between $v$ and
+    $w$, then there must be at least $k$ simple paths of length exactly two between $i(v)$ and
+    $i(w)$.
 \end{proposition}
@@
 
-?? Corollary: $G^d$ graphs.
+Rather than using distance properties directly, \citet{DBLP:conf/cp/McCreeshP15} introduced
+the notion of \emph{supplemental} graphs, with the idea that a constraint programming algorithm can
+search for a mapping which is simultaneously a subgraph isomorphism between several different pairs
+of graphs. Let $G^d$ be the graph with the same set of vertices as $G$, but an edge between vertices
+$v$ and $w$ if the distance between $v$ and $w$ in $G$ is at most $d$.  Similarly, let $G^{n,\ell}$
+be the graph with the same set of vertices as $G$, but an edge between vertices $v$ and $w$ if there
+are at least $n$ simple paths of length exactly $\ell$ between $v$ and $w$ in $G$. Following on from
+\cref{proposition:distances,proposition:paths}, we generalise the result of
+\citeauthor{DBLP:conf/cp/McCreeshP15} as follows.
+
+\begin{corollary}Any homomorphism $i$ from $G$ to $H$ gives a homomorphism $i^d$ from $G^d$ to $H^d$
+    defined by $i^d(v) = i(v)$, for all $d$. Furthermore, if $i$ is locally injective, then for all
+    $n$, there is a homomorphism $i^{n,2}$ from $G^{n,2}$ to
+    $H^{n,2}$ defined by $i^{n,2}(v) = i(v)$.
+\end{corollary}
+
+Note carefully that $i^{n,2}$ is \emph{not} necessarily locally injective, even if $i$ is; we
+illustrate a counter-example in \cref{figure:counterexamples} (bottom right). This is in contrast to
+subgraph isomorphism, where $i^{n,\ell}$ is a subgraph isomorphism for all $\ell$
+\cite{DBLP:conf/cp/McCreeshP15}.
 
@@
     Let $i$ be a homomorphism from $G$ to $H$ where $H$ does
-    not contain any loops. Let $S\subset V$ be such that $G[S]=K_k$. Then $H[i(S)]=K_k$. 
+    not contain any loops. Let $S\subset V$ be such that $G[S]$ is a $k$-vertex clique. Then
+    $H[i(S)]$ is also a $k$-vertex clique.
 \end{proposition}
@@
 
+?? Epimorphism
+
 \section*{Acknowledgements}

### 2021-01-15 paper/paper.tex
Tweaks

@@
 variants. Indeed, this paper proves that many, but not all, of these strategies are also valid in
-the locally injective case, and that simple distance filtering is valid for all
-homomorphisms. However, we also show that \emph{none} of the other strategies are valid for finding
-homomorphisms where there is no injectivity requirement. Finally, we introduce a new filtering
-technique that is based upon finding a maximum clique in the neighbourhood of each pattern vertex, and
-prove that this technique is valid even in the general case. Although this new filtering technique
-involves solving a large number of \NP-complete problems, we demonstrate that it is effective in
-practice, particularly for the non-injective problem where filtering allows a constraint programming
-algorithm to solve instances over eight hundred times faster in aggregate, over a collection of over
-fourteen thousand standard benchmark instances. This shows, for the first time, that modern
-constraint programming techniques can be practical for less constrained graph homomorphism finding
-problems; previous algorithmic approaches have focused instead upon worst-case computational bounds
+the locally injective case, and that simple distance filtering is valid for all homomorphisms.
+However, we also show that \emph{none} of the other strategies are valid for finding homomorphisms
+where there is no injectivity requirement. Finally, we introduce a new filtering technique that is
+based upon finding a maximum clique in the neighbourhood of each pattern vertex, and prove that this
+technique is valid even in the general case. Although this new filtering technique involves solving
+many additional \NP-complete problems as a preprocessing step, we demonstrate that it is effective
+in practice, particularly for the non-injective problem where filtering allows a constraint
+programming algorithm to solve instances over eight hundred times faster in aggregate, over a
+collection of over fourteen thousand standard benchmark instances. This shows, for the first time,
+that modern constraint programming techniques can be practical for less constrained graph
+homomorphism finding problems; previous algorithmic approaches have focused instead upon worst-case
+computational bounds
 \cite{DBLP:journals/csr/FialaK08,DBLP:journals/ipl/Rzazewski14,DBLP:journals/tcs/ChaplickFHPT15},
@@
 
-\begin{proposition}\label{proposition:degreends}
+\begin{proposition}[neighbourhood degree sequences are preserved]\label{proposition:degreends}
     Suppose  $i$ is a locally injective homomorphism from a $G$ to  $H$. Let
@@
 constraints, which are propagated when domains are initialised, before search begins. For each
-domain representing a pattern vertex, any value representing target vertex whose degree or
+domain representing a pattern vertex, any value representing a target vertex whose degree or
 neighbourhood degree sequence is too low may immediately be rejected.
@@
 
-\begin{proposition}\label{proposition:distances}For the problem of finding a subgraph isomorphism
-    $i$ from a graph $G$ to a graph $H$, the following constraints are implied for any pair of
-    vertices $v, w \in \vertexset(G)$:
+\begin{proposition}[paths are preserved by subgraph isomorphisms]\label{proposition:distances}For
+    the problem of finding a subgraph isomorphism $i$ from a graph $G$ to a graph $H$, the following
+    constraints are implied for any pair of vertices $v, w \in \vertexset(G)$:
     \begin{enumerate}
@@
 
-\begin{proposition}\label{proposition:paths}For the problem of finding a locally injective graph
+\begin{proposition}[paths of length two are preserved by locally injective
+    homomorphisms]\label{proposition:paths}For the problem of finding a locally injective graph
     homomorphism $i$ from a graph $G$ to a graph $H$, for any pair of vertices $v, w \in
@@
 between $v$ and $w$. Observe that each $x_j$ is in the neighbourhood of $v$, and so must be mapped
-    to different vertices. Thus each sequence $(i(v), i(x_i), i(w))$ gives a distinct simple path of
+    to different vertices. Thus each sequence $(i(v), i(x_j), i(w))$ gives a distinct simple path of
 length two between $i(v)$ and $i(w)$ as required.\end{proof}

### 2021-01-18 paper/paper.tex
More things

@@
 
-% \usepackage{showframe}
+\usepackage{showframe}
 
@@
 A \emph{graph homomorphism} is a function between two graphs that maps adjacent vertices to adjacent
-vertices. An injective graph homomorphism is more commonly known as a \emph{subgraph isomorphism},
-whilst a \emph{locally injective homomorphism} is one which is injective when restricted to any
-individual vertex and its neighbourhood \cite{DBLP:journals/csr/FialaK08}. Finding any of these three kinds of homomorphism between
-two given graphs is \NP-complete.  The subgraph isomorphism problem arises in many practical
-applications, including ??, and as such there has been a lot of research into finding algorithms
-which perform well empirically. Most approaches are based either upon very fast but simple
-backtracking algorithms,
+vertices. An injective graph homomorphism is more commonly known as a \emph{subgraph isomorphism}, a
+\emph{locally injective homomorphism} is one which is injective when restricted to any individual
+vertex and its neighbourhood \cite{DBLP:journals/csr/FialaK08}, and a homomorphism which is
+surjective on edges is known as an \emph{epimorphism} \cite{DBLP:journals/dam/GayFMSS14}. Finding
+any of these kinds of homomorphism between two given graphs is \NP-complete.  The subgraph
+isomorphism problem arises in many practical applications, including ??, and as such there has been
+a lot of research into finding algorithms which perform well empirically. Most approaches are based
+either upon very fast but simple backtracking algorithms,
 \cite{DBLP:journals/pami/CordellaFSV04,DBLP:journals/bmcbi/BonniciGPSF13,DBLP:conf/gbrpr/CarlettiFSV17}
@@
 where there is no injectivity requirement. Finally, we introduce a new filtering technique that is
-based upon finding a maximum clique in the neighbourhood of each pattern vertex, and prove that this
+based upon finding a maximum clique in the neighbourhood of each domain vertex, and prove that this
 technique is valid even in the general case. Although this new filtering technique involves solving
@@
 
-\paragraph{Degree-based invariants for homomorphisms.} Constraint programming techniques for finding
+\paragraph{Degree-based invariants.} Constraint programming techniques for finding
 homomorphisms can be made much more effective by exploiting certain invariants. It is
@@
 constraints, which are propagated when domains are initialised, before search begins. For each
-domain representing a pattern vertex, any value representing a target vertex whose degree or
+domain representing a domain vertex, any value representing a codomain vertex whose degree or
 neighbourhood degree sequence is too low may immediately be rejected.
@@
 
-\paragraph{Path-based invariants for homomorphisms.} Constraint programming solvers can also exploit
+\paragraph{Path-based invariants.} Constraint programming solvers can also exploit
 invariants that are based upon paths. This is done by automatically adding \emph{implied}
@@
     $n$, there is a homomorphism $i^{n,2}$ from $G^{n,2}$ to
-    $H^{n,2}$ defined by $i^{n,2}(v) = i(v)$.
+    $H^{n,2}$ defined by $i^{n,2}(v) = i(v)$.\label{corollary:lishapes}
 \end{corollary}
@@
 by branching on the variable which has the smallest domain---typically this will be a domain for a
-pattern vertex of high degree, or which has many high degree neighbours. Furthermore, once one such
+domain vertex of high degree, or which has many high degree neighbours. Furthermore, once one such
 guessed assignment has been made, adjacency propagation means many other domains will be
@@
 \cite{DBLP:conf/gg/McCreeshP020}, due to it being the current single strongest solver for the
-subgraph isomorphism problem.
+subgraph isomorphism problem. Adapting the solver to handle locally injective and general
+homomorphism problems required the following straightforward changes.
 
-??
+\paragraph{Injectivity constraints.} For subgraph isomorphism, injectivity is handled by a
+combination of binary constraints and a specialised bit-parallel all-different propagator
+\cite{DBLP:conf/cp/McCreeshP15}. This was disabled entirely for homomorphism problems, and for local
+injectivity only binary constraints were used.
 
-We implemented distance filtering for the basic homomorphism problem, using
-the existing supplemental graphs framework that the solver uses for subgraph isomorphism to create
-pseudo-adjacency constraints for vertices that are distance two away from each other; this was
-routine.
+\paragraph{Path-based filtering.} For subgraph isomorphism, the Glasgow Subgraph Solver
+searches for a mapping which is simultaneously a subgraph isomorphisms between $(G, H)$ and each
+$(G^{n,2}, G^{n,2})$ for each $n$ from $1$ to $4$, and optionally also between $(G^3, H^3)$.
 
-We also implemented clique constraints, initially using the Glasgow Subgraph Solver's existing
-maximum clique solver to perform domain filtering. Preliminary experiments showed that a na{\"\i}ve
-approach, which begins by calculating the maximum clique size for the neighbourhood of each pattern
-vertex and each target vertex, would add as much as three minutes of preprocessing time to some
-problem instances which could otherwise be solved in a few seconds. We therefore use a more
-intelligent approach, which we now describe.
+\paragraph{Degree-based filtering.} This was disabled for homomorphisms and left enabled for locally
+injective homomorphisms, as per \cref{proposition:degreends}. For subgraph isomorphism, the Glasgow
+Subgraph Solver uses degree and neighbourhood degree sequences not just on the original graphs, but
+also on each $(G^{n,\ell}, H^{n,\ell})$ graph pair. This is \emph{not} possible for locally
+injective homomorphisms, due to the counter-example following \ref{corollary:lishapes}.
 
-\subsection{Optimising Maximum Clique Solver Calls}
+\paragraph{Search order heuristics.} The solver's default search order heuristics are based upon
+three principles: that it is good to branch on variables with few remaining values in a domain, that
+it is good to branch on variables corresponding to high-degree vertices, and that it is good to try
+mapping to vertices of high degree
+\cite{DBLP:journals/jair/McCreeshPST18,DBLP:conf/cpaior/ArchibaldDHMP019}. These principles do not
+appear to be specific to subgraph isomorphism, and so we do not alter the search order heuristics.
 
-We calculate the maximum clique size for the neighbourhood of a first pattern vertex $p_1$.  Having
-found a maximum clique $\omega_1$ for a vertex $p$, for every other vertex $q \in \omega_1$, we
-remember that the maximum clique for $q$'s neighbourhood has at least $|\omega_1|$ vertices. We then
-proceed to find a maximum clique $\omega_2$ for a second pattern vertex $p_2$; however, if $p_2$ was
-a member of $\omega_1$, we initialise the clique solver with a lower bound of $n$ rather than
-starting from scratch. We then find the maximum clique size for a third pattern vertex $p_3$, using
-whichever lower bound of $|\omega_1|$ and $|\omega_2|$ is better (if both are valid), and so on.
+\paragraph{Loops.} For the homomorphism decision problem, we modified the solver so that if the
+codomain graph contains a loop, we immediately return a satisfying assignment mapping every vertex
+to the loop.
 
-We then move on to the target vertices, using a similar caching routine. Rather than calculating a
-clique size for every single target vertex, we only calculate a value for target vertices which are
+\subsection{Clique Filtering}
+
+Implementing clique constraints required substantially more work. The Glasgow Subgraph Solver
+contains a maximum clique implementation, which is based upon a variation of
+\citet{DBLP:conf/walcom/TomitaSHTW10}'s MCS algorithm which
+\citet{DBLP:journals/algorithms/Prosser12} calls ``MCSa1'', with bit-parallelism and an altered
+branching scheme for faster optimality proofs \cite{DBLP:conf/cp/McCreeshP14}. The implementation
+also makes use of a shortcut due to \citet{DBLP:journals/jco/BatsynGMP14}, which allows certain
+cliques to be detected without branching---this turns out to be particularly useful in this setting,
+making many of the clique instances generated solvable using only a single greedy colouring step.
+Initially we used this solver to perform domain filtering as a preprocessing step. Preliminary
+experiments showed that a na{\"\i}ve approach, which simply calculates the maximum clique size for
+the neighbourhood of each domain vertex and each codomain vertex, would add as much as three minutes
+of preprocessing time to some problem instances which could otherwise be solved in a few seconds. We
+therefore use a more intelligent approach, which we now describe.
+
+Assuming a problem instance is not detected as obviously unsatisfiable, we calculate the maximum
+clique size for the neighbourhood of every domain vertex in turn. However, having found the maximum
+clique with $k$ vertices in the neighbourhood of a vertex $p$, we remember for every other vertex in
+the maximum clique that its neighbourhood maximum clique size is at least $k$. This can be used to
+accelerate subsequent clique solver calls, by starting the branch and bound algorithm with an
+initial incumbent size of $k$ rather than zero.
+
+We then move on to the codomain vertices, using a similar caching routine. Rather than calculating a
+clique size for every single codomain vertex, we only calculate a value for codomain vertices which are
 present in at least one variable's domain, after other unary constraints have been applied.
 (Depending upon the problem variant, implied constraints like degree and neighbourhood degree
-sequence can eliminate some target vertices from the domains of all pattern variables.)
+sequence can eliminate some codomain vertices from the domains of all domain variables.)
 Additionally, we do not require the maximum clique solver to run to completion and guarantee that it
 has found a maximum clique. Instead, we allow it to stop as soon as it has found a clique with as
-many vertices as the largest pattern clique. This is useful in practice because it may be very hard
-to decide whether a particular target vertex has neighbourhood clique size of, say, 15 or 16, but if
-the largest pattern vertex has a neighbourhood clique size of only 5 then this is irrelevant.
+many vertices as the largest domain clique. This is useful in practice because it may be very hard
+to decide whether a particular codomain vertex has neighbourhood clique size of, say, 15 or 16, but if
+the largest domain vertex has a neighbourhood clique size of only 5 then this is irrelevant.
 
-?? Why this works
+?? Why this works. Clique finding is much easier, colour bounds, etc. Small graphs. ?? See in
+section blah that it's really fast.
 
@@
 
-?? How to log proofs
+Given the complexity of implementations of modern solvers, a critical question is how we can be sure
+that they are producing correct answers. The Glasgow Subgraph Solver's subgraph isomorphism and
+clique algorithms are both \emph{certifying} \cite{DBLP:journals/csr/McConnellMNS11}: that is, they
+can output a mathematical proof that they have reached a correct answer by sound reasoning, and this
+proof log can be verified by the VeriPB proof checking tool
+\cite{DBLP:conf/cp/GochtMMNPT20,DBLP:conf/ijcai/GochtMN20}. Certification involves two stages:
+firstly, the solver translates its inputs into a pseudo-Boolean model (this part is not verified),
+and secondly (for unsatisfiable instances), the solver provides a proof using cutting planes rules
+and reverse unit propagation steps that the pseudo-Boolean model is unsatisfiable.
+
+Implementing proof-logging for the changes not involving clique-finding was routing---although proof
+logging did catch an implementation bug where we were incorrectly applying degree-based filtering on
+the $G^{2,n}$ graphs for locally injective homomorphisms, in violation of the counter-example to
+\cref{corollary:lishapes}. For clique-based filtering, more work was required. We briefly
+outline the steps taken to implement proof logging for the fact that a vertex $p$ in the domain
+graph which is present in a $k$-vertex clique cannot be mapped to a vertex $t$ in the codomain graph
+which is not present in a $k$-vertex clique.
+
+?? Many technicalities.
+
+We have an $x_{p{,}t}$ saying that $p$ gets mapped to $t$. For a regular clique problem, we would have
+binary constraints $\overline{x}_i + \overline{x}_j \ge 1$ for each non-adjacent vertex $i$ and
+$j$. We wish to derive something similar for the non-adjacent vertices in the neighbourhood of $t$.
+However, our constraints are conditional upon $p$ being mapped to $t$, and so instead we start our
+proof step by deriving three sets of conditional\footnote{For the subgraph isomorphism case, these
+constraints need not be conditional.} pseudo-adjacency constraints:
+\begin{itemize}
+    \item $\overline{x}_{p{,}t} + \overline{x}_{q{,}u} + \overline{x}_{r{,}v} \ge 1$
+    \item $\overline{x}_{p{,}t} + \overline{x}_{q{,}u} + \overline{x}_{r{,}u} \ge 1$
+    \item $\overline{x}_{q{,}u} + \overline{x}_{q{,}v} \ge 1$
+\end{itemize}
+Each of these constraints follows by reverse unit propagation from the original model, and so
+requires no further justification in a proof log. We also need an objective constraint. From the
+definition of the homomorphism problem, we have existing adjacency constraints $\overline{x}_{p{,}t}
++ \sum_{u \in \neighbourhood(t)} x_{q{,}u} \ge 1$, and we may sum these constraints for each of the
+$k - 1$ vertices in the $k$-clique that includes $p$ to express that a suitable $k$-clique must
+exist in the codomain graph.
+
+We have now established something that is sufficiently similar to the maximum clique model used by
+?? that the . Colour classes (used for bounds) are in correspondence with colour classes in this
+adapted problem instance, which is similar to microstructure. Produce a proof that there is no
+target clique. The condition appears in the trail, and rather than deriving a contradiction, gets
+$\overline{x}_{p,t} \ge 1$.
+
+This situation is not entirely ideal. The proofs produced this way are inherently dependent upon a
+\emph{particular} choice of $p$ and $t$ vertices. When not proof logging, we need to solve a single
+maximum clique problem for each vertex in the domain, and a restricted clique problem for each
+vertex in the codomain. With proof logging, we must instead solve restricted clique problems for
+each $(p, t)$ assignment that is to be filtered, which is potentially a linear factor blowup. This
+feels morally excessive, since for any given $t$ (and for any clique in the codomain graph which is
+reused) the proofs produced are effectively the same up to a substitution of variable names. We
+would therefore be interested in seeing a proof system extension which would allow us to avoid this
+duplication.
 
@@
     problem variants, with and without clique filtering, and with and without distance filtering for
-    the homomorphism problem. The remaining plots re-display this data, as follows. The three plots
-    on the bottom row zoom in on the cumulative number of instances solved, for the homomorphism
-    problem on the left, the locally injective homomorphism problem in the centre, and the subgraph
-    isomorphism problem on the right. The top centre plot shows the additional number of instances
-    solved at any given time when using the new forms of filtering for all problem variants, and the
-    top right plot shows the aggregate speedups from each form of filtering.}
+    the homomorphism problem; for comparison, results for subgraph isomoprhism using the PathLAD,
+    RI, and VF2 solvers are also shown. The remaining plots re-display this data, as follows. The
+    three plots on the bottom row zoom in on the cumulative number of instances solved, for the
+    homomorphism problem on the left, the locally injective homomorphism problem in the centre, and
+    the subgraph isomorphism problem on the right. The top centre plot shows the additional number
+    of instances solved at any given time when using the new forms of filtering for all problem
+    variants, and the top right plot shows the aggregate speedups from each form of filtering.}
 \end{figure*}
@@
 
-We now evaluate this approach empirically.
+We now evaluate this approach empirically on
 
@@
 ?? Whether cliques are hard
+
+?? Loops
 

### 2021-01-18 paper/paper.tex
Sonja's comments from Overleaf

--- 2021-01-15 paper/paper.tex
@@
 \urlstyle{same}
+\usepackage[textsize=tiny,textwidth=2cm,shadow]{todonotes}
 
+\newcommand{\skcom}[1]{\todo[color=orange!50!white]{{S}: #1}}
+\newcommand{\rcom}[1]{\todo[color=gray!50!white]{{R}: #1}}%reviewers said
 \usepackage{cleveref}
@@
 whilst a \emph{locally injective homomorphism} is one which is injective when restricted to any
-individual vertex and its neighbourhood \cite{DBLP:journals/csr/FialaK08}. Finding any of these three kinds of homomorphism between
-two given graphs is \NP-complete.  The subgraph isomorphism problem arises in many practical
+individual vertex and its neighbourhood \cite{DBLP:journals/csr/FialaK08}.\skcom{It seems a little bit strange to start with a definition.
+How about we describe it more wordy in the introduction?
+Or maybe I would start motivating with subgraph isomorphism(which has an easy definition i words) and the fact that this doesnt ideally match many causes...
+and then go to the definition of general homomorphism?
+} Finding any of these three kinds of homomorphism between
+two given graphs is \NP-complete\skcom{cite a proof of this}.  The subgraph isomorphism problem arises in many practical
 applications, including ??, and as such there has been a lot of research into finding algorithms
@@
 where there is no injectivity requirement. Finally, we introduce a new filtering technique that is
-based upon finding a maximum clique in the neighbourhood of each pattern vertex, and prove that this
-technique is valid even in the general case. Although this new filtering technique involves solving
+based upon finding a maximum clique in the neighbourhood of each pattern vertex, a
+technique, that is valid even in the general case.\skcom{I wouldnt emphasize so much that "we prove" - its a common result and also pretty immediate..} Although this new filtering technique involves solving
 many additional \NP-complete problems as a preprocessing step, we demonstrate that it is effective
@@
 connected by three paths of length two may be mapped onto a pair of vertices connected by a single
-path of length two (see \cref{figure:counterexamples}, top right). Local injectivity is trickier.
-The second property does not hold in full generality, because, for example, a path of length three
+path of length two (see \cref{figure:counterexamples}, top right).
+While standard injectivity prevents any kind of merging, local injectivity prevents two vertices from being merged, only if they are both in the neighbourhood of the same vertex.\skcom{added this}
+So the second property does not hold in full generality, because, for example, a path of length three
 may be mapped onto a triangle whilst respecting local injectivity (\cref{figure:counterexamples},
@@
     $i(w)$.
-\end{proposition}
+\end{proposition}\skcom{I am unsure whether not to write a general statement combining this and proposition 2.2.. but maybe that}
 
@@
     subgraph isomorphisms even if loops are present in the second graph.
-\end{corollary}
+\end{corollary}\skcom{Do you think we should say it or just state in the definition sections that both of these '
+\textit{are} homomorphisms}
 
@@
 proceed to find a maximum clique $\omega_2$ for a second pattern vertex $p_2$; however, if $p_2$ was
-a member of $\omega_1$, we initialise the clique solver with a lower bound of $n$ rather than
+a member of $\omega_1$, we initialise the clique solver with a lower bound of $n$\skcom{why n?} rather than
 starting from scratch. We then find the maximum clique size for a third pattern vertex $p_3$, using
-whichever lower bound of $|\omega_1|$ and $|\omega_2|$ is better (if both are valid), and so on.
+whichever lower bound of $|\omega_1|$ and $|\omega_2|$ is better (if both are valid), and so on.\skcom{I think after you state the first example we can just say that in general, For vertex v, we take the largest lowerbound of maxclique values in its neighbourhood (initialise to 0)?}\skcom{Maybe explicitly say in which order we take the vertices?}
 
@@
 to decide whether a particular target vertex has neighbourhood clique size of, say, 15 or 16, but if
-the largest pattern vertex has a neighbourhood clique size of only 5 then this is irrelevant.
+the largest pattern vertex has a neighbourhood clique size of only 5 then this is irrelevant.\skcom{I would change this bit.
+I think it should be clear that on the second graph we \textit{dont} find a maximum clique, but what we want (for our cliques are preserved constraint) is cliques$\geq$ max size in the pattern vertex or in general of at least the sizes we find in the pattern vertex}
 

### 2021-01-19 paper/paper.tex
More things

--- 2021-01-18 paper/paper.tex
@@
 
-\usepackage{showframe}
+% \usepackage[textsize=tiny,textwidth=2cm,shadow]{todonotes}
+% \newcommand{\skcom}[1]{\todo[color=orange!50!white]{{S}: #1}}
+% \newcommand{\rcom}[1]{\todo[color=gray!50!white]{{R}: #1}}%reviewers said
+% \usepackage{showframe}
 
@@
 
-\author{
-Ciaran McCreesh$^1$
-\and
-Sonja Kraiczy$^2$
-\affiliations
-$^1$University of Glasgow\\
-$^2$University of Oxford\\
-\emails
-ciaran.mccreesh@glasgow.ac.uk
-}
+\author{Paper 4726}
+
+% \author{
+% Ciaran McCreesh$^1$
+% \and
+% Sonja Kraiczy$^2$
+% \affiliations
+% $^1$University of Glasgow\\
+% $^2$University of Oxford\\
+% \emails
+% ciaran.mccreesh@glasgow.ac.uk
+% }
 
@@
 
-A \emph{graph homomorphism} is a function between two graphs that maps adjacent vertices to adjacent
-vertices. An injective graph homomorphism is more commonly known as a \emph{subgraph isomorphism}, a
-\emph{locally injective homomorphism} is one which is injective when restricted to any individual
-vertex and its neighbourhood \cite{DBLP:journals/csr/FialaK08}, and a homomorphism which is
-surjective on edges is known as an \emph{epimorphism} \cite{DBLP:journals/dam/GayFMSS14}. Finding
-any of these kinds of homomorphism between two given graphs is \NP-complete.  The subgraph
-isomorphism problem arises in many practical applications, including ??, and as such there has been
-a lot of research into finding algorithms which perform well empirically. Most approaches are based
-either upon very fast but simple backtracking algorithms,
+The \emph{subgraph isomorphism problem}, as defined by \citet{DBLP:books/fm/GareyJ79}, is to
+determine whether an injective mapping exists from one given graph to another, such that adjacent
+pairs of vertices are mapped to adjacent pairs of vertices, and non-adjacent vertices are mapped to
+non-adjacent vertices.  However, in application-oriented papers, particularly from bioinformatics
+and chemistry, the same name is often implicitly used to mean the \emph{non-induced} version of the
+problem (which does not require that non-adjacency be preserved) since this more accurately models
+the real-world problem being solved \cite{Willett1999,EhrlichR11}. Despite being theoretically hard
+problems, these applications, along with
+others in areas including compilers \cite{DBLP:conf/cp/BlindellLCS15}
+and pattern recognition \cite{DBLP:journals/ijprai/FoggiaPV14},
+have given rise to a large amount of research into designing practical algorithms for solving these
+problems.  Most approaches are based either upon very fast but simple backtracking algorithms
 \cite{DBLP:journals/pami/CordellaFSV04,DBLP:journals/bmcbi/BonniciGPSF13,DBLP:conf/gbrpr/CarlettiFSV17}
@@
 which have higher startup costs but that perform vastly better on harder instances and much more
-consistently on easy instances \cite{DBLP:journals/jair/McCreeshPST18,DBLP:conf/gbrpr/Solnon19}. The
-current state of the art is the Glasgow Subgraph Solver \cite{DBLP:conf/gg/McCreeshP020}, which is a
-dedicated constraint programming solver for subgraph-finding problems. Much of its performance comes
-from inference strategies based upon degrees and neighbourhood degree sequences
+consistently on easy instances \cite{DBLP:journals/jair/McCreeshPST18,DBLP:conf/gbrpr/Solnon19}.
+
+The current state of the art is the Glasgow Subgraph Solver \cite{DBLP:conf/gg/McCreeshP020}, which
+is a dedicated constraint programming solver for subgraph-finding problems. Much of its performance
+comes from inference strategies based upon degrees and neighbourhood degree sequences
 \cite{DBLP:journals/constraints/ZampelliDS10}, counting paths between vertices
@@
 \cite{DBLP:journals/ai/Solnon10}, which can be used to eliminate many infeasible candidate
-assignments without search.
+assignments without search. Like most other solvers, it can handle both the non-induced version of
+the problem, as well as the traditional induced variant, and users can select whichever variant
+better fits their application. It also supports directed edges, and can use an external constraint
+solver to handle other extensions to the basic notion of a graph, such as temporal networks and
+multi-graphs.
 
-One might hope that this inference would also be helpful for other graph homomorphism problem
-variants. Indeed, this paper proves that many, but not all, of these strategies are also valid in
-the locally injective case, and that simple distance filtering is valid for all homomorphisms.
-However, we also show that \emph{none} of the other strategies are valid for finding homomorphisms
-where there is no injectivity requirement. Finally, we introduce a new filtering technique that is
-based upon finding a maximum clique in the neighbourhood of each domain vertex, and prove that this
-technique is valid even in the general case. Although this new filtering technique involves solving
-many additional \NP-complete problems as a preprocessing step, we demonstrate that it is effective
-in practice, particularly for the non-injective problem where filtering allows a constraint
-programming algorithm to solve instances over eight hundred times faster in aggregate, over a
-collection of over fourteen thousand standard benchmark instances. This shows, for the first time,
-that modern constraint programming techniques can be practical for less constrained graph
-homomorphism finding problems; previous algorithmic approaches have focused instead upon worst-case
-computational bounds
-\cite{DBLP:journals/csr/FialaK08,DBLP:journals/ipl/Rzazewski14,DBLP:journals/tcs/ChaplickFHPT15},
+As well as questions over exactly how edges are represented and mapped, some applications would
+prefer not to specify injectivity. A \emph{graph homomorphism} is a function between two graphs that
+maps adjacent vertices to adjacent vertices, with no injectivity requirement, whilst a \emph{locally
+injective homomorphism} is one which is injective when restricted to any individual vertex and its
+neighbourhood. Both are better models for some applications
+\cite{DBLP:journals/dam/FialaKK01,DBLP:conf/semweb/Baget05,DBLP:conf/iccs/CorbyF07,DBLP:journals/csr/FialaK08,DBLP:journals/pvldb/FanLMWW10,DBLP:journals/tcs/ChaplickFHPT15}.
+(There is also an edge-surjective version of the problem known as an \emph{epimorphism}
+\cite{DBLP:journals/dam/GayFMSS14}.)
+
+These injectivity relaxations have received much less attention
+from a practical solving perspective.  One might hope that the inference techniques developed for
+subgraph isomorphism would also be helpful for other graph homomorphism problem variants. Indeed,
+this paper proves that many, but not all, of these strategies are also valid in the locally
+injective case, and that simple distance filtering is valid for all homomorphisms.  However, we also
+show that \emph{none} of the other strategies are valid for finding homomorphisms where there is no
+injectivity requirement.  Finally, we introduce a new filtering technique that is based upon finding
+a maximum clique in the neighbourhood of each domain vertex, which is valid even in the general
+case. Although this new filtering technique involves solving many additional \NP-complete problems
+as a preprocessing step, we demonstrate that it is effective in practice, particularly for the
+non-injective problem where filtering allows a constraint programming algorithm to solve instances
+over eight hundred times faster in aggregate, over a collection of over fourteen thousand standard
+benchmark instances. This shows, for the first time, that modern constraint programming techniques
+can be practical for less constrained graph homomorphism finding problems; previous algorithmic
+approaches have focused instead upon worst-case computational bounds
+(e.g.\
+\citet{DBLP:journals/mst/FominHK07,DBLP:journals/csr/FialaK08,DBLP:journals/ipl/Rzazewski14,DBLP:journals/tcs/ChaplickFHPT15}),
 whose practical utility has yet to be demonstrated. The empirical effectiveness of our results is
@@
 they perform well using off the shelf solvers, rather than because they exactly match ideal domain
-requirements ??cite.
+requirements.
 
@@
 
-We begin by introducing notation and terminology, and by giving the theoretical results which
+We begin by introducing notation and terminology, and by giving the theory which
 support our implementation.
@@
     Suppose  $i$ is a locally injective homomorphism from a $G$ to  $H$. Let
-    $v$ be any vertex. Then:
-    \begin{enumerate}
-        \item $i$ cannot map $v$ to a vertex of lower degree, $\degree_G(v) \le \degree_H(i(v))$.
-        \item Furthermore, $v$ cannot be mapped to any vertex whose neighbourhood degree sequence is
+    $v$ be any vertex. Then $i$ cannot map $v$ to a vertex of lower degree, $\degree_G(v) \le \degree_H(i(v))$.
+    Furthermore, $v$ cannot be mapped to any vertex whose neighbourhood degree sequence is
             not pointwise greater than or equal to its own.
-    \end{enumerate}
 \end{proposition}
@@
 
-However, these invariants do \emph{not} hold for homomorphisms in general. As a simple
+However, these invariants do \emph{not} hold for homomorphisms in general. As a
 counter-example, a homomorphism may map a star graph onto a single edge, as in
@@
 connected by three paths of length two may be mapped onto a pair of vertices connected by a single
-path of length two (see \cref{figure:counterexamples}, top right). Local injectivity is trickier.
-The second property does not hold in full generality, because, for example, a path of length three
-may be mapped onto a triangle whilst respecting local injectivity (\cref{figure:counterexamples},
-bottom left).  However, we observe that a weaker result holds:
+path of length two (see \cref{figure:counterexamples}, top right). While standard injectivity
+prevents any kind of merging, local injectivity prevents two vertices from being merged if they are
+both in the neighbourhood of the same vertex.  So the second property does not hold in full
+generality, because, for example, a path of length three may be mapped onto a triangle whilst
+respecting local injectivity (\cref{figure:counterexamples}, bottom left).  However, we observe that
+a weaker result holds:
 
@@
     to different vertices. Thus each sequence $(i(v), i(x_j), i(w))$ gives a distinct simple path of
-length two between $i(v)$ and $i(w)$ as required.\end{proof}
+length two between $i(v)$ and $i(w)$.\end{proof}
 
@@
     Let $i$ be a homomorphism from $G$ to $H$ where $H$ does
-    not contain any loops. Let $S\subset V$ be such that $G[S]$ is a $k$-vertex clique. Then
+    not contain any loops. Let $S\subseteq V$ be such that $G[S]$ is a $k$-vertex clique. Then
     $H[i(S)]$ is also a $k$-vertex clique.
@@
 
-We will discuss how this is useful in practice in the following section.
+Next we will discuss how this is useful in practice.
 
@@
 these invariants may be used in practice. Our starting point is the Glasgow Subgraph Solver
-\cite{DBLP:conf/gg/McCreeshP020}, due to it being the current single strongest solver for the
-subgraph isomorphism problem. Adapting the solver to handle locally injective and general
-homomorphism problems required the following straightforward changes.
+\cite{DBLP:conf/gg/McCreeshP020}, due to it being the current strongest solver for
+subgraph isomorphism. Adapting the solver to handle locally injective and general homomorphism
+problems required the following straightforward changes.
 
@@
 searches for a mapping which is simultaneously a subgraph isomorphisms between $(G, H)$ and each
-$(G^{n,2}, G^{n,2})$ for each $n$ from $1$ to $4$, and optionally also between $(G^3, H^3)$.
+$(G^{n,2}, G^{n,2})$ for each $n$ from $1$ to $4$, and optionally also between $(G^3, H^3)$. For
+homomorphisms, we use $(G^2, H^2)$, whilst for locally injective homomorphisms we use $(G^{n,2},
+G^{n,2})$ for each $n$ from $1$ to $4$.
 
@@
 also on each $(G^{n,\ell}, H^{n,\ell})$ graph pair. This is \emph{not} possible for locally
-injective homomorphisms, due to the counter-example following \ref{corollary:lishapes}.
+injective homomorphisms, due to the counter-example following \cref{corollary:lishapes}.
 
@@
 present in at least one variable's domain, after other unary constraints have been applied.
-(Depending upon the problem variant, implied constraints like degree and neighbourhood degree
-sequence can eliminate some codomain vertices from the domains of all domain variables.)
 Additionally, we do not require the maximum clique solver to run to completion and guarantee that it
@@
 the largest domain vertex has a neighbourhood clique size of only 5 then this is irrelevant.
-
-?? Why this works. Clique finding is much easier, colour bounds, etc. Small graphs. ?? See in
-section blah that it's really fast.
 
@@
 proof log can be verified by the VeriPB proof checking tool
-\cite{DBLP:conf/cp/GochtMMNPT20,DBLP:conf/ijcai/GochtMN20}. Certification involves two stages:
-firstly, the solver translates its inputs into a pseudo-Boolean model (this part is not verified),
-and secondly (for unsatisfiable instances), the solver provides a proof using cutting planes rules
-and reverse unit propagation steps that the pseudo-Boolean model is unsatisfiable.
+\cite{DBLP:conf/cp/GochtMMNPT20,DBLP:conf/ijcai/GochtMN20}.
 
-Implementing proof-logging for the changes not involving clique-finding was routing---although proof
+Implementing proof-logging for the changes not involving clique-finding was routine---although proof
 logging did catch an implementation bug where we were incorrectly applying degree-based filtering on
 the $G^{2,n}$ graphs for locally injective homomorphisms, in violation of the counter-example to
-\cref{corollary:lishapes}. For clique-based filtering, more work was required. We briefly
-outline the steps taken to implement proof logging for the fact that a vertex $p$ in the domain
-graph which is present in a $k$-vertex clique cannot be mapped to a vertex $t$ in the codomain graph
-which is not present in a $k$-vertex clique.
-
-?? Many technicalities.
-
-We have an $x_{p{,}t}$ saying that $p$ gets mapped to $t$. For a regular clique problem, we would have
-binary constraints $\overline{x}_i + \overline{x}_j \ge 1$ for each non-adjacent vertex $i$ and
-$j$. We wish to derive something similar for the non-adjacent vertices in the neighbourhood of $t$.
-However, our constraints are conditional upon $p$ being mapped to $t$, and so instead we start our
-proof step by deriving three sets of conditional\footnote{For the subgraph isomorphism case, these
-constraints need not be conditional.} pseudo-adjacency constraints:
-\begin{itemize}
-    \item $\overline{x}_{p{,}t} + \overline{x}_{q{,}u} + \overline{x}_{r{,}v} \ge 1$
-    \item $\overline{x}_{p{,}t} + \overline{x}_{q{,}u} + \overline{x}_{r{,}u} \ge 1$
-    \item $\overline{x}_{q{,}u} + \overline{x}_{q{,}v} \ge 1$
-\end{itemize}
-Each of these constraints follows by reverse unit propagation from the original model, and so
-requires no further justification in a proof log. We also need an objective constraint. From the
-definition of the homomorphism problem, we have existing adjacency constraints $\overline{x}_{p{,}t}
-+ \sum_{u \in \neighbourhood(t)} x_{q{,}u} \ge 1$, and we may sum these constraints for each of the
-$k - 1$ vertices in the $k$-clique that includes $p$ to express that a suitable $k$-clique must
-exist in the codomain graph.
-
-We have now established something that is sufficiently similar to the maximum clique model used by
-?? that the . Colour classes (used for bounds) are in correspondence with colour classes in this
-adapted problem instance, which is similar to microstructure. Produce a proof that there is no
-target clique. The condition appears in the trail, and rather than deriving a contradiction, gets
-$\overline{x}_{p,t} \ge 1$.
+\cref{corollary:lishapes}. For clique-based filtering, the process involved substantially more
+technical work. Briefly, for a given domain vertex $v$ that cannot be mapped to a codomain vertex
+$w$, it is possible to instruct the proof verifier to derive a set of conditional clique constraints
+saying that ``either $v$ cannot be mapped to $w$, or the following clique-like problem must be
+satisfiable'' by following steps similar to the maximum common subgraph to clique reduction
+presented by \citet{DBLP:conf/cp/GochtMMNPT20}, and then to reuse clique
+proof-logging techniques to show that this clique-like problem is unsatisfiable.
 
 This situation is not entirely ideal. The proofs produced this way are inherently dependent upon a
-\emph{particular} choice of $p$ and $t$ vertices. When not proof logging, we need to solve a single
-maximum clique problem for each vertex in the domain, and a restricted clique problem for each
-vertex in the codomain. With proof logging, we must instead solve restricted clique problems for
-each $(p, t)$ assignment that is to be filtered, which is potentially a linear factor blowup. This
-feels morally excessive, since for any given $t$ (and for any clique in the codomain graph which is
-reused) the proofs produced are effectively the same up to a substitution of variable names. We
-would therefore be interested in seeing a proof system extension which would allow us to avoid this
-duplication.
+\emph{particular} choice of domain and codomain vertices. When not proof logging, we only need to
+solve a single maximum clique problem for each vertex in the domain, and a restricted clique problem
+for each vertex in the codomain. With proof logging, we must instead solve restricted clique
+problems for each assignment that is to be filtered, which is potentially a linear factor blowup.
+This seems wasteful, since for any given codomain vertex these proofs are effectively the same up to
+substitution of variable names. It would therefore seem desirable to extend the proof system with a
+substitution rule which would allow such proofs to be recycled to avoid duplication.
 
@@
     of instances solved at any given time when using the new forms of filtering for all problem
-    variants, and the top right plot shows the aggregate speedups from each form of filtering.}
+    variants, and the top right plot shows the aggregate speedups from each form of
+    filtering.}\label{figure:plots}
 \end{figure*}
@@
     illustrate that in each the filtering is broadly useful rather than being specific to a single
-    kind of application.}
+    kind of application.}\label{figure:scatters}
 \end{figure*}
 
-We now evaluate this approach empirically on
+We now evaluate this approach empirically. Our experiments are performed on a cluster of machines
+with dual Intel Xeon E5-2697A v4 CPUs with 512GBytes of RAM, running Ubuntu 18.04, and using a
+timeout of one hour. Because injectivity relaxations have received less attention until now, for all
+three problem variants we will use a suite of 14,621 benchmark instances designed for the subgraph
+isomorphism problem \cite{DBLP:conf/lion/KotthoffMS16}. Note that 8,904 of these instances contain
+at least one loop in the codomain graph, and so are trivial for the homomorphism problem---we still
+include these to avoid using different y-axes on different plots.
 
-?? How good is it
+In the top left plot of \cref{figure:plots} we show the cumulative number of instances solved over
+time, for each of the three problem variants, using the Glasgow Subgraph Solver with and without our
+clique-filtering added. For the homomorphism problem, we also compare with distance filtering
+enabled or disabled. For subgraph isomorphism, we also compare against the PathLAD
+\cite{DBLP:conf/lion/KotthoffMS16}, VF2 \cite{DBLP:journals/pami/CordellaFSV04} and RI
+\cite{DBLP:journals/bmcbi/BonniciGPSF13} solvers. Because every Glasgow Subgraph Solver result is so
+much stronger than the other three solvers, it is hard to get any further information from this
+plot. We therefore re-plot zoomed in results on the second row of \cref{figure:plots}, focusing each
+problem variant in turn. The results show that for the homomorphism problem, using either clique or
+distance filtering gives a clear improvement to performance, and using both together is better
+still. For locally injective homomorphisms and subgraph isomorphism, clique filtering also gives an
+improvement.
 
-?? Time on cliques
+To quantify these gains, the top centre plot of \cref{figure:plots} shows the
+additional number of instances solved at any given time (that is, the vertical distance between
+two cumulative plots) when enabling additional filtering. We see that for subgraph isomorphism, at
+the one hour timeout we solve 22 additional instances using clique filtering; for locally injective
+homomorphisms, 21 additional instances; and for homomorphisms, 114, 146 or 193 additional instances
+using just distance filtering, just clique filtering, or both. Alternatively, we can measure the
+horizontal distance between plots for a given choice of timeout $t$ as follows. Let $t'$
+be the last time no greater than $t$ where the slower algorithm solved an instance, and let $u$ be
+the lowest amount of time needed for the faster algorithm to solve the same number of instances.
+Then the aggregate speedup, which we plot in the top right of \cref{figure:plots}, is the ratio of
+$t'$ to $u$ or $u$ to $t'$ as appropriate. At the timeout, we see aggregate speedups of 2.2 for
+locally injective homomorphisms, 3.2 for subgraph isomorphism, and 47.5, 131.7, and 803.5 for
+homomorphisms.
 
-?? Whether cliques are hard
+The crossover point, where clique filtering begins to pay off, is at around the 100ms mark for all
+homomorphism variants, and at around one second for subgraph isomorphism. This does not mean that
+this much time was spent in clique-finding, but rather that this is how long it takes for an
+algorithm with increased filtering to catch up from a slower start. In fact, for domain graphs, no
+maximum clique call took even one millisecond to complete (and none required more than two hundred
+recursive calls from the clique solver), and more time was spent setting up the necessary data
+structures to represent the graphs than on the solving itself. For codomain graphs, 20,838,008 of
+the clique problems we solved took under a millisecond, and the remaining 279 calls took between one
+and four milliseconds, with no instance requiring more than a thousand recursive calls to solve.
+Although it may seem counter-intuitive to try to speed up the solving of a hard problem by first
+solving many other hard problems, these values demonstrate that clique-solving on these smaller
+graphs is \emph{much} easier in practice than solving the larger problem, and so is worthwhile.
 
-?? Loops
+Finally, \cref{figure:scatters}, plots instance-by-instance comparisons of the runtimes. The
+point styles in these plots show different families of benchmark instance, and demonstrate
+that the benefits of filtering are not restricted to one application.
 
-\section{Conclusion and Future Work}
+\section{Conclusion}
 
-?? Neighbourhood colourability? Cores?
+These results show that constraint programming approaches are viable for a wider range of subgraph
+mapping finding problems than had previously been considered, which will increase the versatility of
+subgraph solvers for application users going forwards. We also saw that using clique-finding as a
+filtering step could give moderate to very large speedups for different problem variants---although
+only through careful engineering choices, to mitigate the potential cost of solving \NP-hard
+problems. We would be interested to see if filtering based upon other hard problems could give
+similar benefits---for example, homomorphisms also interact in potentially helpful ways with various
+colouring properties.
 
-?? Epimorphism
-
-\section*{Acknowledgements}
-
-This work was supported by the Engineering and Physical Sciences Research Council [grant number
-EP/P026842/1].
+% \section*{Acknowledgements}
+% 
+% This work was supported by the Engineering and Physical Sciences Research Council [grant number
+% EP/P026842/1].
 

### 2021-01-19 paper/paper.tex
Last tweaks

@@
 assignments without search. Like most other solvers, it can handle both the non-induced version of
-the problem, as well as the traditional induced variant, and users can select whichever variant
+the problem and the traditional induced variant, and users can select whichever variant
 better fits their application. It also supports directed edges, and can use an external constraint
@@
 therefore follows that for the homomorphism \emph{decision} problem where there are no injectivity
-constraints, any problem instance which has a loop in the codomain graph is trivially satisfiable,
-although the \emph{counting} problem remains \#\P-hard \cite{DBLP:journals/rsa/DyerG00}.
+constraints, any problem instance which has a loop in the codomain graph is trivially satisfiable
+(although the \emph{counting} problem remains \#\P-hard \cite{DBLP:journals/rsa/DyerG00}).
 
@@
 connected by three paths of length two may be mapped onto a pair of vertices connected by a single
-path of length two (see \cref{figure:counterexamples}, top right). While standard injectivity
-prevents any kind of merging, local injectivity prevents two vertices from being merged if they are
-both in the neighbourhood of the same vertex.  So the second property does not hold in full
-generality, because, for example, a path of length three may be mapped onto a triangle whilst
-respecting local injectivity (\cref{figure:counterexamples}, bottom left).  However, we observe that
-a weaker result holds:
+path of length two (see \cref{figure:counterexamples}, top right). The second property also does not
+hold in full generality for local injectivity, because, for example, a path of length three may be
+mapped onto a triangle (\cref{figure:counterexamples}, bottom left): while standard injectivity
+prevents any kind of merging, local injectivity only prevents two vertices from being merged if they
+are both in the neighbourhood of the same vertex. However, we observe that a weaker result does
+hold:
 
@@
 
-Rather than using distance properties directly, \citet{DBLP:conf/cp/McCreeshP15} introduced
-the notion of \emph{supplemental} graphs, with the idea that a constraint programming algorithm can
-search for a mapping which is simultaneously a subgraph isomorphism between several different pairs
-of graphs. Let $G^d$ be the graph with the same set of vertices as $G$, but an edge between vertices
-$v$ and $w$ if the distance between $v$ and $w$ in $G$ is at most $d$.  Similarly, let $G^{n,\ell}$
-be the graph with the same set of vertices as $G$, but an edge between vertices $v$ and $w$ if there
-are at least $n$ simple paths of length exactly $\ell$ between $v$ and $w$ in $G$. Following on from
-\cref{proposition:distances,proposition:paths}, we generalise the result of
-\citeauthor{DBLP:conf/cp/McCreeshP15} as follows.
+Rather than using distance properties directly as \citet{DBLP:conf/cp/AudemardLMGP14} did,
+\citet{DBLP:conf/cp/McCreeshP15} introduced the notion of \emph{supplemental} graphs, with the idea
+that a constraint programming algorithm can search for a mapping which is simultaneously a subgraph
+isomorphism between several different pairs of graphs. Let $G^d$ be the graph with the same set of
+vertices as $G$, but an edge between vertices $v$ and $w$ if the distance between $v$ and $w$ in $G$
+is at most $d$.  Similarly, let $G^{n,\ell}$ be the graph with the same set of vertices as $G$, but
+an edge between vertices $v$ and $w$ if there are at least $n$ simple paths of length exactly $\ell$
+between $v$ and $w$ in $G$. Following on from \cref{proposition:distances,proposition:paths}, we
+generalise the result of \citeauthor{DBLP:conf/cp/McCreeshP15} as follows.
 
@@
 
-\paragraph{Cliques and homomorphisms.} Finally, we come to the main result of this paper. Observe
-that until now, we have not seen any unary constraints which are valid for the general homomorphism
+\paragraph{Cliques and homomorphisms.} We now discuss the main new technique of this paper. Observe
+that so far, we have not seen any unary constraints which are valid for the general homomorphism
 problem. All existing constraint programming approaches for the subgraph isomorphism problem begin
 by branching on the variable which has the smallest domain---typically this will be a domain for a
-domain vertex of high degree, or which has many high degree neighbours. Furthermore, once one such
+vertex of high degree, or which has many high degree neighbours. Furthermore, once one such
 guessed assignment has been made, adjacency propagation means many other domains will be
@@
 initially be of the same size, which will make it much harder for a solver to know where to start.
-This motivates the following observation.
+This motivates the following.
 
@@
 codomain graph contains a loop, we immediately return a satisfying assignment mapping every vertex
-to the loop.
+to one of these loops.
 
@@
 \citet{DBLP:conf/walcom/TomitaSHTW10}'s MCS algorithm which
-\citet{DBLP:journals/algorithms/Prosser12} calls ``MCSa1'', with bit-parallelism and an altered
+\citet{DBLP:journals/algorithms/Prosser12} calls ``MCSa1'', with bit-parallelism
+\cite{DBLP:journals/ol/SegundoMRH13} and an altered
 branching scheme for faster optimality proofs \cite{DBLP:conf/cp/McCreeshP14}. The implementation
@@
 cliques to be detected without branching---this turns out to be particularly useful in this setting,
-making many of the clique instances generated solvable using only a single greedy colouring step.
+making many of the clique instances generated solvable without branching.
 Initially we used this solver to perform domain filtering as a preprocessing step. Preliminary
-experiments showed that a na{\"\i}ve approach, which simply calculates the maximum clique size for
+experiments showed that a na{\"\i}ve approach, which calculates the maximum clique size for
 the neighbourhood of each domain vertex and each codomain vertex, would add as much as three minutes
 of preprocessing time to some problem instances which could otherwise be solved in a few seconds. We
-therefore use a more intelligent approach, which we now describe.
+therefore invested more engineering effort, as follows.
 
 Assuming a problem instance is not detected as obviously unsatisfiable, we calculate the maximum
-clique size for the neighbourhood of every domain vertex in turn. However, having found the maximum
+clique size for the neighbourhood of every domain vertex in turn. However, having found a maximum
 clique with $k$ vertices in the neighbourhood of a vertex $p$, we remember for every other vertex in
-the maximum clique that its neighbourhood maximum clique size is at least $k$. This can be used to
+this maximum clique that its neighbourhood maximum clique size is at least $k$. This can be used to
 accelerate subsequent clique solver calls, by starting the branch and bound algorithm with an
@@
 
-Given the complexity of implementations of modern solvers, a critical question is how we can be sure
+Given the complexity of modern solvers, a critical question is how we can be sure
 that they are producing correct answers. The Glasgow Subgraph Solver's subgraph isomorphism and
@@
 clique-filtering added. For the homomorphism problem, we also compare with distance filtering
-enabled or disabled. For subgraph isomorphism, we also compare against the PathLAD
+enabled or disabled. For subgraph isomorphism, we further compare against the PathLAD
 \cite{DBLP:conf/lion/KotthoffMS16}, VF2 \cite{DBLP:journals/pami/CordellaFSV04} and RI
-\cite{DBLP:journals/bmcbi/BonniciGPSF13} solvers. Because every Glasgow Subgraph Solver result is so
-much stronger than the other three solvers, it is hard to get any further information from this
-plot. We therefore re-plot zoomed in results on the second row of \cref{figure:plots}, focusing each
-problem variant in turn. The results show that for the homomorphism problem, using either clique or
+\cite{DBLP:journals/bmcbi/BonniciGPSF13} solvers. To make the details of this plot readable, the
+second row of \cref{figure:plots} zooms in on the results for each
+problem variant in turn. We see that for the homomorphism problem, using either clique or
 distance filtering gives a clear improvement to performance, and using both together is better
-still. For locally injective homomorphisms and subgraph isomorphism, clique filtering also gives an
-improvement.
+still. For both injective problems, clique filtering gives a more moderate improvement.
 
@@
 the lowest amount of time needed for the faster algorithm to solve the same number of instances.
-Then the aggregate speedup, which we plot in the top right of \cref{figure:plots}, is the ratio of
-$t'$ to $u$ or $u$ to $t'$ as appropriate. At the timeout, we see aggregate speedups of 2.2 for
-locally injective homomorphisms, 3.2 for subgraph isomorphism, and 47.5, 131.7, and 803.5 for
+The \emph{aggregate speedup}, which we plot in the top right of \cref{figure:plots}, is the ratio of
+$t'$ to $u$ or $u$ to $t'$ as appropriate. At the one hour timeout, we see aggregate speedups of 2.2
+for locally injective homomorphisms, 3.2 for subgraph isomorphism, and 47.5, 131.7, and 803.5 for
 homomorphisms.
 
-The crossover point, where clique filtering begins to pay off, is at around the 100ms mark for all
-homomorphism variants, and at around one second for subgraph isomorphism. This does not mean that
-this much time was spent in clique-finding, but rather that this is how long it takes for an
-algorithm with increased filtering to catch up from a slower start. In fact, for domain graphs, no
-maximum clique call took even one millisecond to complete (and none required more than two hundred
-recursive calls from the clique solver), and more time was spent setting up the necessary data
-structures to represent the graphs than on the solving itself. For codomain graphs, 20,838,008 of
-the clique problems we solved took under a millisecond, and the remaining 279 calls took between one
-and four milliseconds, with no instance requiring more than a thousand recursive calls to solve.
-Although it may seem counter-intuitive to try to speed up the solving of a hard problem by first
-solving many other hard problems, these values demonstrate that clique-solving on these smaller
-graphs is \emph{much} easier in practice than solving the larger problem, and so is worthwhile.
+The crossover point, where clique filtering begins to pay off, is at around the 100ms mark without
+injectivity, and at around one second for the injective problems. This does not show how much time
+was spent in clique-finding, but rather how long it takes for an algorithm with increased
+filtering to catch up from a slower start. In fact, for domain graphs, no maximum clique call took
+even one millisecond to complete (and none required more than two hundred recursive calls from the
+clique solver), and more time was spent setting up the data structures to represent the
+neighbourhood graphs than on the actual solving. For codomain graphs, 20,838,008 of the clique
+problems we solved took under a millisecond, and the remaining 279 calls took between one and four
+milliseconds, with no instance requiring more than a thousand recursive calls.  Although it may seem
+counter-intuitive to try to speed up the solving of a hard problem by first solving many other hard
+problems, these values demonstrate that clique-solving on these smaller graphs is \emph{much} easier
+in practice than solving the larger problem, and so is worthwhile.
 
-Finally, \cref{figure:scatters}, plots instance-by-instance comparisons of the runtimes. The
+Finally, \cref{figure:scatters} plots instance-by-instance comparisons of the runtimes. The
 point styles in these plots show different families of benchmark instance, and demonstrate
-that the benefits of filtering are not restricted to one application.
+that the benefits of filtering are not restricted to one application. They also show that clique
+filtering very rarely makes performance a lot worse, but often makes it a lot better.
 
@@
 
-These results show that constraint programming approaches are viable for a wider range of subgraph
-mapping finding problems than had previously been considered, which will increase the versatility of
-subgraph solvers for application users going forwards. We also saw that using clique-finding as a
+Our results have shown that constraint programming approaches are viable for a wider range of subgraph
+mapping finding problems than had previously been considered, and will increase the versatility of
+subgraph solvers for application users going forwards. We saw that using clique-finding as a
 filtering step could give moderate to very large speedups for different problem variants---although

### 2021-01-20 glasgow-subgraph-solver/src/cheap_all_different.cc
Import GSS code

@@ -1,0 +1,119 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "cheap_all_different.hh"
+
+#include <tuple>
+#include <type_traits>
+
+using std::conditional_t;
+using std::tuple;
+using std::shared_ptr;
+using std::vector;
+
+namespace
+{
+    template <bool proof_>
+    auto cheap_all_different_with_optional_proofs(
+            unsigned target_size,
+            vector<HomomorphismDomain> & domains,
+            const shared_ptr<Proof> & proof) -> bool
+    {
+        // Pick domains smallest first; ties are broken by smallest .v first.
+        // For each count p we have a linked list, whose first member is
+        // first[p].  The element following x in one of these lists is next[x].
+        // Any domain with a count greater than domains.size() is put
+        // int the "count==domains.size()" bucket.
+        // The "first" array is sized to be able to hold domains.size()+1
+        // elements
+        vector<int> first(target_size + 1, -1), next(target_size, -1);
+
+        [[ maybe_unused ]] conditional_t<proof_, vector<int>, tuple<> > lhs, hall_lhs, hall_rhs;
+
+        // Iterate backwards, because we insert elements at the head of
+        // lists and we want the sort to be stable
+        for (int i = int(domains.size()) - 1 ; i >= 0; --i) {
+            unsigned count = domains.at(i).count;
+            if (count > domains.size())
+                count = domains.size();
+            next.at(i) = first.at(count);
+            first.at(count) = i;
+        }
+
+        // counting all-different
+        SVOBitset domains_so_far{ target_size, 0 }, hall{ target_size, 0 };
+        unsigned neighbours_so_far = 0;
+
+        [[ maybe_unused ]] conditional_t<proof_, unsigned, tuple<> > last_outputted_hall_size{};
+
+        for (unsigned i = 0 ; i <= domains.size() ; ++i) {
+            // iterate over linked lists
+            int domain_index = first[i];
+            while (domain_index != -1) {
+                auto & d = domains.at(domain_index);
+
+                if constexpr (proof_)
+                    lhs.push_back(d.v);
+
+                [[ maybe_unused ]] conditional_t<proof_, unsigned, tuple<> > old_d_values_count;
+                if constexpr (proof_)
+                    old_d_values_count = d.values.count();
+
+                d.values.intersect_with_complement(hall);
+                d.count = d.values.count();
+
+                if constexpr (proof_)
[…]

### 2021-01-20 glasgow-subgraph-solver/src/clique.cc

@@ -1,0 +1,650 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "clique.hh"
+#include "watches.hh"
+#include "svo_bitset.hh"
+#include "proof.hh"
+#include "configuration.hh"
+
+#include <algorithm>
+#include <list>
+#include <numeric>
+#include <random>
+#include <tuple>
+#include <type_traits>
+#include <utility>
+#include <vector>
+
+using std::conditional_t;
+using std::find;
+using std::iota;
+using std::is_same;
+using std::list;
+using std::make_tuple;
+using std::mt19937;
+using std::move;
+using std::pair;
+using std::reverse;
+using std::sort;
+using std::string_view;
+using std::swap;
+using std::to_string;
+using std::vector;
+
+namespace
+{
+    enum class SearchResult
+    {
+        Aborted,
+        Restart,
+        Complete,
+        DecidedTrue
+    };
+
+    struct Incumbent
+    {
+        unsigned value = 0;
+        vector<int> c;
+
+        auto update(const vector<int> & new_c, unsigned long long & find_nodes, unsigned long long & prove_nodes) -> void
+        {
+            if (new_c.size() > value) {
+                find_nodes += prove_nodes;
+                prove_nodes = 0;
+                value = new_c.size();
+                c = new_c;
+            }
+        }
+    };
+
+    template <typename EntryType_>
+    struct FlatWatchTable
+    {
+        vector<EntryType_> data;
+
+        EntryType_ & operator[] (int x)
+        {
+            return data[x];
+        }
+    };
+
+    struct CliqueRunner
+    {
+        const CliqueParams & params;
+        Incumbent incumbent;
+
+        int size;
+        vector<SVOBitset> adj, connected_table;
+        vector<int> order, invorder;
+
+        Watches<int, FlatWatchTable> watches;
+
+        mt19937 global_rand;
+
+        int * space;
+
+        CliqueRunner(const InputGraph & g, const CliqueParams & p) :
+            params(p),
+            size(g.size()),
+            adj(g.size(), SVOBitset{ unsigned(size), 0 }),
+            order(size),
+            invorder(size),
+            space(nullptr)
+        {
+            space = new int[size * (size + 1) * 2];
+
+            if (params.restarts_schedule->might_restart())
+                watches.table.data.resize(g.size());
+
+            // populate our order with every vertex initially
+            iota(order.begin(), order.end(), 0);
+
[…]

### 2021-01-20 glasgow-subgraph-solver/src/common_subgraph.cc

@@ -1,0 +1,471 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "common_subgraph.hh"
+#include "proof.hh"
+#include "configuration.hh"
+#include "clique.hh"
+#include "proof.hh"
+
+#include <algorithm>
+#include <map>
+#include <optional>
+#include <set>
+#include <string_view>
+#include <utility>
+#include <vector>
+
+using std::function;
+using std::make_optional;
+using std::make_unique;
+using std::map;
+using std::min;
+using std::nullopt;
+using std::optional;
+using std::pair;
+using std::set;
+using std::string;
+using std::string_view;
+using std::tuple;
+using std::vector;
+
+namespace
+{
+    enum class SearchResult
+    {
+        Aborted,
+        Complete,
+        DecidedTrue,
+        SatisfiableButKeepGoing
+    };
+
+    struct SplitDomains
+    {
+        vector<pair<set<int>, set<int> > > partitions;
+    };
+
+    struct Assignments
+    {
+        vector<pair<int, int> > assigned;
+        vector<int> rejected;
+    };
+
+    auto assignments_as_proof_decisions(const Assignments & assignments) -> vector<pair<int, int> >
+    {
+        vector<pair<int, int> > trail;
+        for (auto & [ l, r ] : assignments.assigned)
+            trail.emplace_back(l, r);
+        return trail;
+    }
+
+    struct CommonSubgraphRunner
+    {
+        const InputGraph & first;
+        const InputGraph & second;
+        const CommonSubgraphParams & params;
+
+        CommonSubgraphRunner(const InputGraph & f, const InputGraph & s, const CommonSubgraphParams & p) :
+            first(f),
+            second(s),
+            params(p)
+        {
+        }
+
+        auto branch_assigning(const SplitDomains & d, int left, int right) -> SplitDomains
+        {
+            SplitDomains result;
+
+            for (auto & [ l, r ] : d.partitions) {
+                map<tuple<bool, bool, string_view, string_view>, pair<set<int>, set<int> > > new_partitions;
+
+                string no_label;
+                auto partition_of = [&] (const InputGraph & g, int w, int v) -> tuple<bool, bool, string_view, string_view> {
+                    return tuple{
+                        g.adjacent(w, v),
+                        g.adjacent(v, w),
+                            g.adjacent(w, v) ? g.edge_label(w, v) : no_label,
+                            g.adjacent(v, w) ? g.edge_label(v, w) : no_label };
+                };
+
[…]

### 2021-01-20 glasgow-subgraph-solver/src/configuration.cc

@@ -1,0 +1,16 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "configuration.hh"
+
+using std::string;
+
+UnsupportedConfiguration::UnsupportedConfiguration(const string & message) noexcept :
+    _what(message)
+{
+}
+
+auto UnsupportedConfiguration::what() const noexcept -> const char *
+{
+    return _what.c_str();
+}
+

### 2021-01-20 glasgow-subgraph-solver/src/formats/csv.cc

@@ -1,0 +1,133 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "formats/csv.hh"
+#include "formats/input_graph.hh"
+
+#include <fstream>
+#include <optional>
+#include <tuple>
+#include <unordered_map>
+#include <vector>
+
+using std::ifstream;
+using std::nullopt;
+using std::optional;
+using std::string;
+using std::tuple;
+using std::unordered_map;
+using std::vector;
+
+namespace
+{
+    auto read_csv(ifstream && infile, const string & filename, const optional<unordered_map<string, string> > & rename_map) -> InputGraph
+    {
+        if (! infile)
+            throw GraphFileError{ filename, "error opening file", false };
+
+        unordered_map<string, int> vertices;
+        unordered_map<string, string> vertex_labels;
+        vector<tuple<int, int, string> > edges;
+        bool seen_vertex_label = false, seen_edge_label = false, seen_directed_edge = false;
+
+        string line;
+
+        while (getline(infile, line)) {
+            auto pos = line.find_first_of(",>");
+            if (string::npos == pos)
+                throw GraphFileError{ filename, "expected a comma but didn't get one", true };
+
+            string left = line.substr(0, pos), right = line.substr(pos + 1), label;
+            char delim = line.at(pos);
+
+            auto pos2 = right.find(',');
+            if (string::npos != pos2) {
+                label = right.substr(pos2 + 1);
+                right = right.substr(0, pos2);
+            }
+
+            if (right.empty() && ! left.empty()) {
+                if (! label.empty()) {
+                    seen_vertex_label = true;
+                    vertex_labels.emplace(left, label);
+                }
+
+                vertices.emplace(left, vertices.size());
+            }
+            else {
+                int left_idx = vertices.emplace(left, vertices.size()).first->second;
+                int right_idx = vertices.emplace(right, vertices.size()).first->second;
+
+                if (! label.empty())
+                    seen_edge_label = true;
+
+                if (delim == '>') {
+                    seen_directed_edge = true;
+                    edges.emplace_back(left_idx, right_idx, label);
+                }
+                else {
+                    edges.emplace_back(left_idx, right_idx, label);
+                    edges.emplace_back(right_idx, left_idx, label);
+                }
+            }
+        }
+
+        InputGraph result{ int(vertices.size()), seen_vertex_label, seen_edge_label };
+
+        for (auto & [ f, t, l ] : edges)
[…]

### 2021-01-20 glasgow-subgraph-solver/src/formats/dimacs.cc

@@ -1,0 +1,65 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "formats/dimacs.hh"
+#include "formats/input_graph.hh"
+#include "formats/graph_file_error.hh"
+
+#include <fstream>
+#include <regex>
+
+using std::getline;
+using std::ifstream;
+using std::regex;
+using std::smatch;
+using std::stoi;
+using std::string;
+using std::to_string;
+
+auto read_dimacs(ifstream && infile, const string & filename) -> InputGraph
+{
+    InputGraph result{ 0, false, false };
+
+    string line;
+    while (getline(infile, line)) {
+        if (line.empty())
+            continue;
+
+        /* Lines are comments, a problem description (contains the number of
+         * vertices), or an edge. */
+        static const regex
+            comment{ R"(c(\s.*)?)" },
+            problem{ R"(p\s+(edge|col)\s+(\d+)\s+(\d+)?\s*)" },
+            edge{ R"(e\s+(\d+)\s+(\d+)\s*)" };
+
+        smatch match;
+        if (regex_match(line, match, comment)) {
+            /* Comment, ignore */
+        }
+        else if (regex_match(line, match, problem)) {
+            /* Problem. Specifies the size of the graph. Must happen exactly
+             * once. */
+            if (0 != result.size())
+                throw GraphFileError{ filename, "multiple 'p' lines encountered", true };
+            result.resize(stoi(match.str(2)));
+        }
+        else if (regex_match(line, match, edge)) {
+            /* An edge. DIMACS files are 1-indexed. We assume we've already had
+             * a problem line (if not our size will be 0, so we'll throw). */
+            int a{ stoi(match.str(1)) }, b{ stoi(match.str(2)) };
+            if (0 == a || 0 == b || a > result.size() || b > result.size())
+                throw GraphFileError{ filename, "line '" + line + "' edge index out of bounds", true };
+            result.add_edge(a - 1, b - 1);
+        }
+        else
+            throw GraphFileError{ filename, "cannot parse line '" + line + "'", true };
+    }
+
+    if (! infile.eof())
+        throw GraphFileError{ filename, "error reading file", true };
+
+    for (int v = 0 ; v < result.size() ; ++v)
+        result.set_vertex_name(v, to_string(v + 1));
+
+    return result;
+}
+

### 2021-01-20 glasgow-subgraph-solver/src/formats/graph_file_error.cc

@@ -1,0 +1,25 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "formats/graph_file_error.hh"
+#include "formats/input_graph.hh"
+
+#include <fstream>
+
+using std::string;
+
+GraphFileError::GraphFileError(const string & filename, const string & message, bool x) noexcept :
+    _what("Error reading graph file '" + filename + "': " + message),
+    _exists(x)
+{
+}
+
+auto GraphFileError::what() const noexcept -> const char *
+{
+    return _what.c_str();
+}
+
+auto GraphFileError::file_at_least_existed() const noexcept -> bool
+{
+    return _exists;
+}
+

### 2021-01-20 glasgow-subgraph-solver/src/formats/input_graph.cc

@@ -1,0 +1,168 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "formats/input_graph.hh"
+
+#include <algorithm>
+#include <cstdint>
+#include <limits>
+#include <string>
+#include <type_traits>
+#include <vector>
+
+#include <boost/bimap.hpp>
+#include <boost/bimap/unordered_set_of.hpp>
+
+using std::distance;
+using std::find;
+using std::function;
+using std::numeric_limits;
+using std::make_optional;
+using std::make_pair;
+using std::map;
+using std::max;
+using std::move;
+using std::nullopt;
+using std::optional;
+using std::pair;
+using std::string;
+using std::string_view;
+using std::to_string;
+using std::vector;
+
+using Names = boost::bimaps::bimap<boost::bimaps::unordered_set_of<int>, boost::bimaps::unordered_set_of<string> >;
+
+struct InputGraph::Imp
+{
+    int size = 0;
+    bool has_vertex_labels, has_edge_labels;
+    map<pair<int, int>, string> edges;
+    vector<string> vertex_labels;
+    Names vertex_names;
+    bool loopy = false, directed = false;
+};
+
+InputGraph::InputGraph(int size, bool v, bool e) :
+    _imp(new Imp{ })
+{
+    _imp->has_vertex_labels = v;
+    _imp->has_edge_labels = e;
+    _imp->loopy = false;
+    _imp->directed = false;
+
+    if (0 != size)
+        resize(size);
+}
+
+InputGraph::~InputGraph() = default;
+
+InputGraph::InputGraph(InputGraph && other) = default;
+
+auto InputGraph::resize(int size) -> void
+{
+    _imp->size = size;
+    _imp->vertex_labels.resize(size);
+}
+
+auto InputGraph::add_edge(int a, int b) -> void
+{
+    _imp->edges.emplace(make_pair(a, b), "");
+    _imp->edges.emplace(make_pair(b, a), "");
+    if (a == b)
+        _imp->loopy = true;
+}
+
+auto InputGraph::add_directed_edge(int a, int b, string_view label) -> void
+{
+    _imp->directed = true;
+
+    _imp->edges.emplace(make_pair(a, b), label).first->second = label;
+    if (a == b)
+        _imp->loopy = true;
+}
+
+auto InputGraph::adjacent(int a, int b) const -> bool
+{
+    return _imp->edges.count({ a, b });
+}
+
[…]

### 2021-01-20 glasgow-subgraph-solver/src/formats/lad.cc

@@ -1,0 +1,97 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "formats/lad.hh"
+#include "formats/input_graph.hh"
+
+#include <fstream>
+#include <map>
+
+using std::ifstream;
+using std::map;
+using std::pair;
+using std::string;
+using std::to_string;
+
+namespace
+{
+    auto read_word(ifstream & infile) -> int
+    {
+        int x;
+        infile >> x;
+        return x;
+    }
+
+    auto read_any_lad(ifstream && infile, const string & filename,
+            bool directed,
+            bool vertex_labels,
+            bool edge_labels) -> InputGraph
+    {
+        InputGraph result{ 0, vertex_labels, edge_labels };
+
+        result.resize(read_word(infile));
+        if (! infile)
+            throw GraphFileError{ filename, "error reading size", true };
+
+        for (int r = 0 ; r < result.size() ; ++r) {
+            if (vertex_labels) {
+                int l = read_word(infile);
+                if (! infile)
+                    throw GraphFileError{ filename, "error reading label", true };
+
+                result.set_vertex_label(r, to_string(l));
+            }
+
+            int c_end = read_word(infile);
+            if (! infile)
+                throw GraphFileError{ filename, "error reading edges count", true };
+
+            for (int c = 0 ; c < c_end ; ++c) {
+                int e = read_word(infile);
+
+                if (e < 0 || e >= result.size())
+                    throw GraphFileError{ filename, "edge index out of bounds", true };
+
+                if (edge_labels) {
+                    int l = read_word(infile);
+                    if (l < 0)
+                        throw GraphFileError{ filename, "edge label invalid", true };
+
+                    result.add_directed_edge(r, e, to_string(l));
+                }
+                else if (directed)
+                    result.add_directed_edge(r, e, "");
+                else
+                    result.add_edge(r, e);
+            }
+        }
+
+        string rest;
+        if (infile >> rest)
+            throw GraphFileError{ filename, "EOF not reached, next text is \"" + rest + "\"", true };
+        if (! infile.eof())
+            throw GraphFileError{ filename, "EOF not reached", true };
+
+        return result;
+    }
+}
+
+auto read_lad(ifstream && infile, const string & filename) -> InputGraph
+{
+    return read_any_lad(move(infile), filename, false, false, false);
+}
+
[…]

### 2021-01-20 glasgow-subgraph-solver/src/formats/read_file_format.cc

@@ -1,0 +1,132 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "formats/read_file_format.hh"
+#include "formats/dimacs.hh"
+#include "formats/lad.hh"
+#include "formats/csv.hh"
+#include "formats/vfmcs.hh"
+
+#include <fstream>
+#include <regex>
+#include <sstream>
+#include <vector>
+
+using std::ifstream;
+using std::ios;
+using std::move;
+using std::regex;
+using std::smatch;
+using std::stoi;
+using std::string;
+using std::stringstream;
+using std::to_string;
+using std::vector;
+
+auto detect_format(ifstream & infile, const string & filename) -> string
+{
+    string line;
+    if (! getline(infile, line) || line.empty())
+        throw GraphFileError{ filename, "unable to read file to detect file format", true };
+
+    static const regex
+        dimacs_comment{ R"(c(\s.*)?)" },
+        dimacs_problem{ R"(p\s+(edge|col)\s+(\d+)\s+(\d+)?\s*)" },
+        lad_header{ R"(\d+)" },
+        lad_zero_labelled_line{ R"(0 \d+)" },
+        lad_zero_unlabelled_line{ R"(0)" },
+        lad_line{ R"(\d+\s+(\d+\s+)*\d+\s*)" },
+        csv_problem{ R"(\S+,\S+)" };
+
+    smatch match;
+    if (regex_match(line, match, dimacs_comment)) {
+        while (regex_match(line, match, dimacs_comment)) {
+            // looks like a DIMACS comment, ignore
+            if (! getline(infile, line) || line.empty())
+                throw GraphFileError{ filename, "unable to auto-detect file format (entirely c lines?)", true };
+        }
+        if (! regex_match(line, match, dimacs_problem))
+            throw GraphFileError{ filename, "unable to auto-detect file format (c line not followed by a p line?)", true };
+        return "dimacs";
+    }
+    else if (regex_match(line, match, dimacs_problem))
+        return "dimacs";
+    else if (regex_match(line, match, lad_header)) {
+        if ("0" == line)
+            return "lad";
+
+        // got to figure out whether we're labelled or not
+        if (! getline(infile, line) || line.empty())
+            throw GraphFileError{ filename, "unable to auto-detect file format (number followed by nothing)", true };
+        if (regex_match(line, match, lad_zero_labelled_line))
+            return "labelledlad";
[…]

### 2021-01-20 glasgow-subgraph-solver/src/glasgow_clique_solver.cc

@@ -1,0 +1,223 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "formats/read_file_format.hh"
+#include "clique.hh"
+#include "configuration.hh"
+#include "proof.hh"
+
+#include <boost/program_options.hpp>
+
+#include <chrono>
+#include <cstdlib>
+#include <ctime>
+#include <exception>
+#include <iomanip>
+#include <iostream>
+#include <memory>
+#include <optional>
+
+#include <unistd.h>
+
+namespace po = boost::program_options;
+
+using std::boolalpha;
+using std::cerr;
+using std::cout;
+using std::endl;
+using std::exception;
+using std::function;
+using std::localtime;
+using std::make_optional;
+using std::make_pair;
+using std::make_shared;
+using std::make_unique;
+using std::put_time;
+using std::string;
+using std::string_view;
+
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+using std::chrono::operator""s;
+using std::chrono::seconds;
+using std::chrono::steady_clock;
+using std::chrono::system_clock;
+
+auto colour_class_order_from_string(string_view s) -> ColourClassOrder
+{
+    if (s == "colour")
+        return ColourClassOrder::ColourOrder;
+    else if (s == "singletons-first")
+        return ColourClassOrder::SingletonsFirst;
+    else if (s == "sorted")
+        return ColourClassOrder::Sorted;
+    else
+        throw UnsupportedConfiguration{ "Unknown colour class order '" + string(s) + "'" };
+}
+
+auto main(int argc, char * argv[]) -> int
+{
+    try {
+        po::options_description display_options{ "Program options" };
+        display_options.add_options()
+            ("help",                                         "Display help information")
+            ("timeout",            po::value<int>(),         "Abort after this many seconds")
+            ("format",             po::value<string>(),      "Specify input file format (auto, lad, labelledlad, dimacs)")
+            ("decide",             po::value<int>(),         "Solve this decision problem");
+
+        po::options_description configuration_options{ "Advanced configuration options" };
+        configuration_options.add_options()
+            ("colour-ordering",    po::value<string>(),      "Specify colour-ordering (colour / singletons-first / sorted)")
+            ("input-order",                                  "Use the input order for colouring (usually a bad idea)")
+            ("restarts-constant",  po::value<int>(),         "How often to perform restarts (disabled by default)")
[…]

### 2021-01-20 glasgow-subgraph-solver/src/glasgow_common_subgraph_solver.cc

@@ -1,0 +1,216 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "formats/read_file_format.hh"
+#include "common_subgraph.hh"
+#include "configuration.hh"
+#include "proof.hh"
+
+#include <boost/program_options.hpp>
+
+#include <chrono>
+#include <cstdlib>
+#include <ctime>
+#include <exception>
+#include <iomanip>
+#include <iostream>
+#include <memory>
+#include <optional>
+
+#include <unistd.h>
+
+namespace po = boost::program_options;
+
+using std::boolalpha;
+using std::cerr;
+using std::cout;
+using std::endl;
+using std::exception;
+using std::function;
+using std::localtime;
+using std::make_optional;
+using std::make_pair;
+using std::make_shared;
+using std::make_unique;
+using std::put_time;
+using std::string;
+using std::string_view;
+
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+using std::chrono::operator""s;
+using std::chrono::seconds;
+using std::chrono::steady_clock;
+using std::chrono::system_clock;
+
+auto main(int argc, char * argv[]) -> int
+{
+    try {
+        po::options_description display_options{ "Program options" };
+        display_options.add_options()
+            ("help",                                         "Display help information")
+            ("timeout",            po::value<int>(),         "Abort after this many seconds")
+            ("decide",             po::value<int>(),         "Solve this decision problem")
+            ("count-solutions",                              "Count the number of solutions (--decide only)")
+            ("print-all-solutions",                          "Print out every solution, rather than one (--decide only)")
+            ("connected",                                    "Only find connected graphs")
+            ("clique",                                       "Use the clique solver")
+            ;
+
+        po::options_description input_options{ "Input file options" };
+        input_options.add_options()
+            ("format",             po::value<string>(),      "Specify input file format (auto, lad, vertexlabelledlad, labelledlad, dimacs)")
+            ("first-format",       po::value<string>(),      "Specify input file format just for the first graph")
+            ("second-format",      po::value<string>(),      "Specify input file format just for the second graph");
+        display_options.add(input_options);
+
+        po::options_description proof_logging_options{ "Proof logging options" };
+        proof_logging_options.add_options()
+            ("prove",               po::value<string>(),       "Write unsat proofs to this filename (suffixed with .opb and .veripb)")
[…]

### 2021-01-20 glasgow-subgraph-solver/src/glasgow_subgraph_solver.cc

@@ -1,0 +1,440 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "formats/read_file_format.hh"
+#include "homomorphism.hh"
+#include "sip_decomposer.hh"
+#include "lackey.hh"
+#include "symmetries.hh"
+#include "restarts.hh"
+#include "verify.hh"
+#include "proof.hh"
+
+#include <boost/program_options.hpp>
+
+#include <chrono>
+#include <cstdlib>
+#include <ctime>
+#include <exception>
+#include <iomanip>
+#include <iostream>
+#include <memory>
+#include <vector>
+
+#include <unistd.h>
+
+namespace po = boost::program_options;
+
+using std::boolalpha;
+using std::cerr;
+using std::cout;
+using std::endl;
+using std::exception;
+using std::function;
+using std::list;
+using std::localtime;
+using std::make_pair;
+using std::make_shared;
+using std::make_unique;
+using std::pair;
+using std::put_time;
+using std::string;
+using std::vector;
+
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+using std::chrono::operator""s;
+using std::chrono::seconds;
+using std::chrono::steady_clock;
+using std::chrono::system_clock;
+
+auto main(int argc, char * argv[]) -> int
+{
+    try {
+        po::options_description display_options{ "Program options" };
+        display_options.add_options()
+            ("help",                                         "Display help information")
+            ("timeout",            po::value<int>(),         "Abort after this many seconds")
+            ("parallel",                                     "Use auto-configured parallel search (highly nondeterministic runtimes)");
+
+        po::options_description problem_options{ "Problem options" };
+        problem_options.add_options()
+            ("noninjective",                                 "Drop the injectivity requirement")
+            ("locally-injective",                            "Require only local injectivity")
+            ("count-solutions",                              "Count the number of solutions")
+            ("print-all-solutions",                          "Print out every solution, rather than one")
+            ("induced",                                      "Find an induced mapping");
+        display_options.add(problem_options);
+
+        po::options_description input_options{ "Input file options" };
+        input_options.add_options()
+            ("format",             po::value<string>(),      "Specify input file format (auto, lad, vertexlabelledlad, labelledlad, dimacs)")
+            ("pattern-format",     po::value<string>(),      "Specify input file format just for the pattern graph")
+            ("target-format",      po::value<string>(),      "Specify input file format just for the target graph");
+        display_options.add(input_options);
+
[…]

### 2021-01-20 glasgow-subgraph-solver/src/homomorphism.cc

@@ -1,0 +1,550 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "homomorphism.hh"
+#include "clique.hh"
+#include "configuration.hh"
+#include "graph_traits.hh"
+#include "homomorphism_domain.hh"
+#include "homomorphism_model.hh"
+#include "homomorphism_searcher.hh"
+#include "homomorphism_traits.hh"
+#include "thread_utils.hh"
+#include "proof.hh"
+
+#include <algorithm>
+#include <atomic>
+#include <condition_variable>
+#include <limits>
+#include <map>
+#include <memory>
+#include <mutex>
+#include <optional>
+#include <thread>
+#include <unordered_set>
+#include <utility>
+
+#include <boost/thread/barrier.hpp>
+#include <boost/functional/hash.hpp>
+
+using std::atomic;
+using std::function;
+using std::make_optional;
+using std::make_unique;
+using std::map;
+using std::move;
+using std::mutex;
+using std::optional;
+using std::pair;
+using std::size_t;
+using std::sort;
+using std::string;
+using std::thread;
+using std::to_string;
+using std::unique_lock;
+using std::unique_ptr;
+using std::unordered_set;
+using std::vector;
+
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+using std::chrono::steady_clock;
+using std::chrono::operator""ms;
+
+using boost::barrier;
+using boost::hash_combine;
+
+namespace
+{
+    struct HomomorphismSolver
+    {
+        using Domains = vector<HomomorphismDomain>;
+
+        const HomomorphismModel & model;
+        const HomomorphismParams & params;
+
+        HomomorphismSolver(const HomomorphismModel & m, const HomomorphismParams & p) :
+            model(m),
+            params(p)
+        {
+        }
+    };
+
+    struct SequentialSolver :
+        HomomorphismSolver
+    {
+        using HomomorphismSolver::HomomorphismSolver;
+
+        auto solve() -> HomomorphismResult
+        {
+            HomomorphismResult result;
+
+            // domains
+            Domains domains(model.pattern_size, HomomorphismDomain{ model.target_size });
+            if (! model.initialise_domains(domains)) {
+                result.complete = true;
+                model.add_extra_stats(result.extra_stats);
+                return result;
+            }
+
+            // assignments
+            HomomorphismAssignments assignments;
+            assignments.values.reserve(model.pattern_size);
+
+            // start search timer
[…]

### 2021-01-20 glasgow-subgraph-solver/src/homomorphism_model.cc

@@ -1,0 +1,1076 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "homomorphism_model.hh"
+#include "homomorphism_traits.hh"
+#include "configuration.hh"
+#include "clique.hh"
+
+#include <chrono>
+#include <functional>
+#include <list>
+#include <map>
+#include <set>
+#include <sstream>
+#include <string>
+#include <utility>
+#include <vector>
+
+using std::greater;
+using std::list;
+using std::make_optional;
+using std::make_unique;
+using std::map;
+using std::max;
+using std::nullopt;
+using std::optional;
+using std::pair;
+using std::set;
+using std::shared_ptr;
+using std::string;
+using std::string_view;
+using std::stringstream;
+using std::to_string;
+using std::vector;
+
+using std::chrono::duration_cast;
+using std::chrono::milliseconds;
+using std::chrono::steady_clock;
+
+namespace
+{
+    auto calculate_n_shape_graphs(const HomomorphismParams & params) -> unsigned
+    {
+        return 1 +
+            (supports_exact_path_graphs(params) ? params.number_of_exact_path_graphs : 0) +
+            (supports_distance2_graphs(params) ? 1 : 0) +
+            (supports_distance3_graphs(params) ? 1 : 0) +
+            (supports_k4_graphs(params) ? 1 : 0);
+    }
+
+    auto find_clique(
+            const shared_ptr<Timeout> & timeout,
+            unsigned size,
+            const vector<SVOBitset> & rows,
+            unsigned g,
+            unsigned max_graphs,
+            unsigned v,
+            optional<int> largest_if_target,
+            vector<int> & best_knowns,
+            list<string> & build_times,
+            list<string> & solve_times,
+            list<string> & find_nodes,
+            list<string> & prove_nodes) -> int
+    {
+        if (largest_if_target && (best_knowns[v] >= *largest_if_target))
+            return best_knowns[v];
+
+        auto start_time = steady_clock::now();
+
+        CliqueParams params;
+        params.timeout = timeout;
+        params.start_time = steady_clock::now();
+        params.decide = nullopt;
+        params.restarts_schedule = make_unique<NoRestartsSchedule>();
+        if (largest_if_target)
+            params.stop_after_finding = *largest_if_target - 1;
+
+        vector<int> include(size, -1), invinclude(size, 0);
[…]

### 2021-01-20 glasgow-subgraph-solver/src/homomorphism_searcher.cc

@@ -1,0 +1,664 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "homomorphism_searcher.hh"
+#include "cheap_all_different.hh"
+
+using std::max;
+using std::move;
+using std::mt19937;
+using std::numeric_limits;
+using std::pair;
+using std::string;
+using std::swap;
+using std::to_string;
+using std::uniform_int_distribution;
+using std::vector;
+
+HomomorphismSearcher::HomomorphismSearcher(const HomomorphismModel & m, const HomomorphismParams & p,
+        const DuplicateSolutionFilterer & d) :
+    model(m),
+    params(p),
+    _duplicate_solution_filterer(d)
+{
+    if (might_have_watches(params)) {
+        watches.table.target_size = model.target_size;
+        watches.table.data.resize(model.pattern_size * model.target_size);
+    }
+}
+
+auto HomomorphismSearcher::assignments_as_proof_decisions(const HomomorphismAssignments & assignments) const -> vector<pair<int, int> >
+{
+    vector<pair<int, int> > trail;
+    for (auto & a : assignments.values)
+        if (a.is_decision)
+            trail.emplace_back(a.assignment.pattern_vertex, a.assignment.target_vertex);
+    return trail;
+}
+
+auto HomomorphismSearcher::solution_in_proof_form(const HomomorphismAssignments & assignments) const -> vector<pair<NamedVertex, NamedVertex> >
+{
+    vector<pair<NamedVertex, NamedVertex> > solution;
+    for (auto & a : assignments.values)
+        if (solution.end() == find_if(solution.begin(), solution.end(),
+                    [&] (const auto & t) { return unsigned(t.first.first) == a.assignment.pattern_vertex; }))
+            solution.emplace_back(
+                    model.pattern_vertex_for_proof(a.assignment.pattern_vertex),
+                    model.target_vertex_for_proof(a.assignment.target_vertex));
+    return solution;
+}
+
+auto HomomorphismSearcher::expand_to_full_result(const HomomorphismAssignments & assignments, VertexToVertexMapping & mapping) -> void
+{
+    for (auto & a : assignments.values)
+        mapping.emplace(a.assignment.pattern_vertex, a.assignment.target_vertex);
+}
+
+auto HomomorphismSearcher::save_result(const HomomorphismAssignments & assignments, HomomorphismResult & result) -> void
+{
+    expand_to_full_result(assignments, result.mapping);
+
[…]

### 2021-01-20 glasgow-subgraph-solver/src/homomorphism_traits.cc

@@ -1,0 +1,55 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "homomorphism_traits.hh"
+
+auto supports_exact_path_graphs(const HomomorphismParams & params) -> bool
+{
+    return (! params.no_supplementals) && (params.injectivity != Injectivity::NonInjective);
+}
+
+auto supports_distance2_graphs(const HomomorphismParams & params) -> bool
+{
+    // exact path graphs are better
+    return (! params.no_supplementals) && (! supports_exact_path_graphs(params));
+}
+
+auto supports_k4_graphs(const HomomorphismParams & params) -> bool
+{
+    return (! params.no_supplementals) && params.k4 && (params.injectivity != Injectivity::NonInjective);
+}
+
+auto supports_distance3_graphs(const HomomorphismParams & params) -> bool
+{
+    return (! params.no_supplementals) && params.distance3 && (params.injectivity == Injectivity::Injective);
+}
+
+auto might_have_watches(const HomomorphismParams & params) -> bool
+{
+    return params.restarts_schedule->might_restart();
+}
+
+auto is_nonshrinking(const HomomorphismParams & params) -> bool
+{
+    return params.injectivity == Injectivity::Injective;
+}
+
+auto degree_and_nds_are_preserved(const HomomorphismParams & params) -> bool
+{
+    return params.injectivity != Injectivity::NonInjective;
+}
+
+auto degree_and_nds_are_exact(const HomomorphismParams & params, unsigned pattern_size, unsigned target_size) -> bool
+{
+    return params.induced && pattern_size == target_size;
+}
+
+auto global_degree_is_preserved(const HomomorphismParams & params) -> bool
+{
+    return params.injectivity == Injectivity::Injective;
+}
+
+auto can_use_clique(const HomomorphismParams & params) -> bool
+{
+    return (! params.count_solutions) && (! params.lackey) && params.clique_detection && (! params.proof);
+}
+

### 2021-01-20 glasgow-subgraph-solver/src/lackey.cc

@@ -1,0 +1,216 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "lackey.hh"
+
+#include <fstream>
+#include <map>
+#include <mutex>
+
+using std::endl;
+using std::function;
+using std::ifstream;
+using std::map;
+using std::mutex;
+using std::ofstream;
+using std::string;
+using std::to_string;
+using std::unique_lock;
+
+DisobedientLackeyError::DisobedientLackeyError(const std::string & m) noexcept :
+    _what(m)
+{
+}
+
+auto DisobedientLackeyError::what() const throw () -> const char *
+{
+    return _what.c_str();
+}
+
+struct Lackey::Imp
+{
+    mutex external_solver_mutex;
+
+    ofstream send_to;
+    ifstream read_from;
+    const InputGraph & pattern_graph;
+    const InputGraph & target_graph;
+
+    long number_of_checks = 0, number_of_propagations = 0, number_of_deletions = 0, number_of_calls = 0;
+};
+
+Lackey::Lackey(const string & send_to_name, const string & read_from_name,
+        const InputGraph & pattern_graph, const InputGraph & target_graph) :
+    _imp(new Imp{ {}, ofstream{ send_to_name }, ifstream{ read_from_name }, pattern_graph, target_graph })
+{
+    if ((! _imp->read_from) || (! _imp->send_to))
+        throw DisobedientLackeyError{ "error setting up lackey communication using " + send_to_name + " and " + read_from_name };
+}
+
+Lackey::~Lackey()
+{
+    if (_imp->send_to) {
+        _imp->send_to << "Q 0" << endl;
+    }
+}
+
+auto Lackey::check_solution(
+        const VertexToVertexMapping & m,
+        bool partial,
+        bool all_solutions,
+        const function<auto (int, int) -> bool> & deletion) -> bool
+{
+    unique_lock<mutex> lock{ _imp->external_solver_mutex };
+    ++_imp->number_of_calls;
+
+    string command;
+    if (partial) {
+        if (deletion) {
+            ++_imp->number_of_propagations;
+            command = "P";
+        }
+        else {
+            ++_imp->number_of_checks;
+            command = "C";
+        }
+    }
+    else {
+        ++_imp->number_of_checks;
+        if (all_solutions)
+            command = "A";
+        else
+            command = "F";
+    }
+
+    _imp->send_to << command << " " << m.size();
[…]

### 2021-01-20 glasgow-subgraph-solver/src/plot_glasgow_solver_proofs.cc

@@ -1,0 +1,331 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include <boost/program_options.hpp>
+
+#include <cstdlib>
+#include <exception>
+#include <fstream>
+#include <iomanip>
+#include <iostream>
+#include <map>
+#include <memory>
+#include <sstream>
+#include <string>
+#include <vector>
+
+namespace po = boost::program_options;
+
+using std::cerr;
+using std::cout;
+using std::endl;
+using std::exception;
+using std::ifstream;
+using std::make_unique;
+using std::map;
+using std::ofstream;
+using std::string;
+using std::stringstream;
+using std::unique_ptr;
+using std::vector;
+
+auto read_instance_file(const string & instance_result_file_name, const string & prefix,
+        map<string, string> & keys, bool & aborted, bool & excluded) -> void
+{
+    ifstream result_file{ instance_result_file_name };
+    if (! result_file) {
+        cerr << "Error reading " << instance_result_file_name << endl;
+        throw 0;
+    }
+    string key_line;
+    while (getline(result_file, key_line)) {
+        auto pos = key_line.find('=');
+        if (string::npos == pos) {
+            cerr << "Couldn't parse '" << key_line << "' in '" << instance_result_file_name << endl;
+            throw 0;
+        }
+
+        auto k = key_line.substr(0, pos);
+        k.erase(k.find_last_not_of(" ") + 1, k.length());
+        auto v = key_line.substr(pos + 1);
+        v.erase(0, v.find_first_not_of(" "));
+        keys.emplace(prefix + k, v);
+
+        if (k == "status") {
+            if (v == "true") {
+            }
+            else if (v == "false") {
+            }
+            else if (v == "aborted" || v == "tooslow") {
+                aborted = true;
+            }
+            else if (v == "excluded") {
+                excluded = true;
+            }
+            else {
+                cerr << "Couldn't parse status value '" << v << "' in " << instance_result_file_name << endl;
+                throw 0;
+            }
+        }
+    }
+}
+
+auto read_veripb_file(const string & veripb_file_name, const string & prefix,
+        map<string, string> & keys, bool & aborted, bool & failed) -> void
+{
+    ifstream result_file{ veripb_file_name };
+    if (! result_file) {
+        cerr << "Error reading " << veripb_file_name << endl;
+        throw 0;
+    }
+    string key_line;
+
+    while (getline(result_file, key_line)) {
[…]

### 2021-01-20 glasgow-subgraph-solver/src/proof.cc

@@ -1,0 +1,981 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "proof.hh"
+
+#include <algorithm>
+#include <iterator>
+#include <fstream>
+#include <map>
+#include <memory>
+#include <sstream>
+#include <tuple>
+#include <utility>
+
+#include <boost/iostreams/device/file.hpp>
+#include <boost/iostreams/filter/bzip2.hpp>
+#include <boost/iostreams/filtering_stream.hpp>
+#include <boost/iostreams/stream.hpp>
+
+using std::copy;
+using std::endl;
+using std::find;
+using std::function;
+using std::istreambuf_iterator;
+using std::make_unique;
+using std::map;
+using std::max;
+using std::min;
+using std::move;
+using std::ofstream;
+using std::optional;
+using std::ostream;
+using std::ostreambuf_iterator;
+using std::pair;
+using std::set;
+using std::string;
+using std::stringstream;
+using std::to_string;
+using std::tuple;
+using std::unique_ptr;
+using std::vector;
+
+using boost::iostreams::bzip2_compressor;
+using boost::iostreams::file_sink;
+using boost::iostreams::filtering_ostream;
+
+namespace
+{
+    auto make_compressed_ostream(const string & fn) -> unique_ptr<ostream>
+    {
+        auto out = make_unique<filtering_ostream>();
+        out->push(bzip2_compressor());
+        out->push(file_sink(fn));
+        return out;
+    }
+}
+
+ProofError::ProofError(const string & m) noexcept :
+    _message("Proof error: " + m)
+{
+}
+
+auto ProofError::what() const noexcept -> const char *
+{
+    return _message.c_str();
+}
+
+struct Proof::Imp
+{
+    string opb_filename, log_filename;
+    stringstream model_stream, model_prelude_stream;
+    unique_ptr<ostream> proof_stream;
+    bool friendly_names;
+    bool bz2 = false;
+
+    map<pair<long, long>, string> variable_mappings;
+    map<long, string> binary_variable_mappings;
+    map<tuple<long, long, long>, string> connected_variable_mappings;
+    map<tuple<long, long, long, long>, string> connected_variable_mappings_aux;
+    map<long, long> at_least_one_value_constraints, at_most_one_value_constraints, injectivity_constraints;
[…]

### 2021-01-20 glasgow-subgraph-solver/src/restarts.cc

@@ -1,0 +1,174 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "restarts.hh"
+#include <algorithm>
+#include <cmath>
+
+using std::advance;
+using std::distance;
+using std::round;
+
+using std::chrono::milliseconds;
+using std::chrono::steady_clock;
+
+auto NoRestartsSchedule::did_a_backtrack() -> void
+{
+}
+
+auto NoRestartsSchedule::did_a_restart() -> void
+{
+}
+
+auto NoRestartsSchedule::should_restart() -> bool
+{
+    return false;
+}
+
+auto NoRestartsSchedule::might_restart() -> bool
+{
+    return false;
+}
+
+auto NoRestartsSchedule::clone() -> NoRestartsSchedule *
+{
+    return new NoRestartsSchedule(*this);
+}
+
+LubyRestartsSchedule::LubyRestartsSchedule(long long m) :
+    _backtracks_remaining(m)
+{
+    _current_sequence = _sequence.insert(_sequence.end(), m);
+}
+
+LubyRestartsSchedule::LubyRestartsSchedule(const LubyRestartsSchedule & other) :
+    _backtracks_remaining(other._backtracks_remaining),
+    _sequence(other._sequence),
+    _current_sequence(_sequence.begin())
+{
+    advance(_current_sequence, distance(other._sequence.begin(), other._current_sequence));
+}
+
+auto LubyRestartsSchedule::did_a_backtrack() -> void
+{
+    --_backtracks_remaining;
+}
+
+auto LubyRestartsSchedule::did_a_restart() -> void
+{
+    if (next(_current_sequence) == _sequence.end()) {
+        _sequence.insert(_sequence.end(), _sequence.begin(), _sequence.end());
+        _sequence.push_back(_sequence.back() * 2);
+    }
+    ++_current_sequence;
+    _backtracks_remaining = *_current_sequence;
+}
+
+auto LubyRestartsSchedule::should_restart() -> bool
+{
+    return _backtracks_remaining <= 0;
+}
+
+auto LubyRestartsSchedule::clone() -> LubyRestartsSchedule *
+{
+    return new LubyRestartsSchedule(*this);
+}
+
+auto LubyRestartsSchedule::might_restart() -> bool
+{
+    return true;
+}
+
+GeometricRestartsSchedule::GeometricRestartsSchedule(double v, double m) :
[…]

### 2021-01-20 glasgow-subgraph-solver/src/sip_to_opb.cc

@@ -1,0 +1,219 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "formats/read_file_format.hh"
+
+#include <boost/program_options.hpp>
+
+#include <algorithm>
+#include <functional>
+#include <iomanip>
+#include <iostream>
+#include <iterator>
+#include <map>
+#include <sstream>
+#include <string>
+#include <vector>
+
+#include <unistd.h>
+
+namespace po = boost::program_options;
+
+using std::cerr;
+using std::copy;
+using std::cout;
+using std::endl;
+using std::exception;
+using std::greater;
+using std::istreambuf_iterator;
+using std::map;
+using std::ostreambuf_iterator;
+using std::pair;
+using std::sort;
+using std::string;
+using std::string_view;
+using std::stringstream;
+using std::vector;
+
+auto main(int argc, char * argv[]) -> int
+{
+    try {
+        po::options_description display_options{ "Program options" };
+        display_options.add_options()
+            ("help",                                         "Display help information")
+            ("format",             po::value<string>(),      "Specify input file format (auto, lad, vertexlabelledlad, labelledlad, dimacs)")
+            ("degree",                                       "Apply degree preprocessing")
+            ("nds",                                          "Also apply neighbourhood degree sequence preprocessing")
+            ;
+
+        po::options_description all_options{ "All options" };
+        all_options.add_options()
+            ("pattern-file", "Specify the pattern file")
+            ("target-file",  "Specify the target file")
+            ;
+        all_options.add(display_options);
+
+        po::positional_options_description positional_options;
+        positional_options
+            .add("pattern-file", 1)
+            .add("target-file", 1)
+            ;
+
+        po::variables_map options_vars;
+        po::store(po::command_line_parser(argc, argv)
+                .options(all_options)
+                .positional(positional_options)
+                .run(), options_vars);
+        po::notify(options_vars);
+
+        /* --help? Show a message, and exit. */
+        if (options_vars.count("help")) {
+            cout << "Usage: " << argv[0] << " [options] pattern target" << endl;
+            cout << endl;
+            cout << display_options << endl;
+            return EXIT_SUCCESS;
+        }
+
+        /* No algorithm or no input file specified? Show a message and exit. */
[…]

### 2021-01-20 glasgow-subgraph-solver/src/symmetries.cc

@@ -1,0 +1,172 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "symmetries.hh"
+#include "config.hh"
+
+#include <istream>
+#include <ostream>
+#include <sstream>
+#include <string>
+#include <vector>
+
+#if defined(STD_FS_IS_EXPERIMENTAL)
+#  include <experimental/filesystem>
+#elif defined(STD_FS_IS_STD)
+#  include <filesystem>
+#elif defined(STD_FS_IS_BOOST)
+#  include <boost/filesystem.hpp>
+#endif
+
+#include <fcntl.h>
+#include <sys/types.h>
+#include <sys/wait.h>
+#include <unistd.h>
+
+using std::endl;
+using std::getline;
+using std::pair;
+using std::string;
+using std::stringstream;
+using std::to_string;
+using std::vector;
+
+#if defined(STD_FS_IS_EXPERIMENTAL)
+using std::experimental::filesystem::exists;
+using std::experimental::filesystem::path;
+#elif defined(STD_FS_IS_STD)
+using std::filesystem::exists;
+using std::filesystem::path;
+#elif defined(STD_FS_IS_BOOST)
+using boost::filesystem::exists;
+using boost::filesystem::path;
+#endif
+
+GapFailedUs::GapFailedUs(const std::string & message) noexcept :
+    _what("Running 'gap for symmetry detection failed: " + message)
+{
+}
+
+auto GapFailedUs::what() const noexcept -> const char *
+{
+    return _what.c_str();
+}
+
+auto find_symmetries(
+        const char * const argv0,
+        const InputGraph & graph,
+        std::list<std::pair<std::string, std::string> > & constraints,
+        std::string & size) -> void
+{
+    path gap_helper_file_path = argv0;
+    if (gap_helper_file_path.has_filename()) {
+        gap_helper_file_path = gap_helper_file_path.parent_path() / "gap" / "findDPfactorsOfGraphs.g";
+    }
+
+    if (! exists(gap_helper_file_path))
+        throw GapFailedUs{ "couldn't find gap/findDPfactorsOfGraphs.g, which we need for symmetry detection" };
+
+#if defined(STD_FS_IS_BOOST)
+    string gap_helper_file_path_str = gap_helper_file_path.string();
+#else
+    string gap_helper_file_path_str = gap_helper_file_path;
+#endif
+
+    stringstream stdin_stream;
[…]

### 2021-01-20 glasgow-subgraph-solver/src/timeout.cc

@@ -1,0 +1,92 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "timeout.hh"
+
+#include <atomic>
+#include <condition_variable>
+#include <mutex>
+#include <thread>
+
+#include <iostream>
+#include <iomanip>
+#include <ctime>
+#include <chrono>
+
+using std::atomic;
+using std::condition_variable;
+using std::cv_status;
+using std::make_unique;
+using std::mutex;
+using std::thread;
+using std::unique_lock;
+
+using std::chrono::operator""s;
+using std::chrono::seconds;
+using std::chrono::system_clock;
+
+struct Timeout::Detail
+{
+    bool aborted = false;
+    thread timeout_thread;
+    mutex timeout_mutex;
+    condition_variable timeout_cv;
+    atomic<bool> abort;
+};
+
+Timeout::Timeout(const seconds limit) :
+    _detail(make_unique<Detail>())
+{
+    _detail->abort.store(false);
+    if (0s != limit) {
+        _detail->timeout_thread = thread([limit, &detail = this->_detail] {
+                auto abort_time = system_clock::now() + limit;
+                {
+                    /* Sleep until either we've reached the time limit,
+                     * or we've finished all the work. */
+                    unique_lock<mutex> guard(detail->timeout_mutex);
+                    while (! detail->abort.load()) {
+                        if (cv_status::timeout == detail->timeout_cv.wait_until(guard, abort_time)) {
+                            /* We've woken up, and it's due to a timeout. */
+                            detail->aborted = true;
+                            break;
+                        }
+                    }
+                }
+                detail->abort.store(true);
+                });
+    }
+}
+
+Timeout::~Timeout()
+{
+    stop();
+}
+
+auto Timeout::should_abort() const -> bool
+{
+    return _detail->abort.load();
+}
+
+auto Timeout::aborted() const -> bool
+{
+    return _detail->aborted;
+}
+
+auto Timeout::trigger_early_abort() -> void
+{
+    return _detail->abort.store(true);
+}
+
+auto Timeout::stop() -> void
+{
+    /* Clean up the timeout thread */
+    if (_detail->timeout_thread.joinable()) {
+        {
+            unique_lock<mutex> guard(_detail->timeout_mutex);
+            _detail->abort.store(true);
+            _detail->timeout_cv.notify_all();
+        }
+        _detail->timeout_thread.join();
+    }
+}
+

### 2021-01-20 glasgow-subgraph-solver/src/verify.cc

@@ -1,0 +1,100 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "verify.hh"
+#include <map>
+
+using std::map;
+using std::string;
+
+BuggySolution::BuggySolution(const string & message) noexcept :
+    _what(message)
+{
+}
+
+auto BuggySolution::what() const noexcept -> const char *
+{
+    return _what.c_str();
+}
+
+auto verify_homomorphism(
+        const InputGraph & pattern,
+        const InputGraph & target,
+        bool injective,
+        bool locally_injective,
+        bool induced,
+        const map<int, int> & mapping) -> void
+{
+    // nothing to verify, if unsat
+    if (mapping.empty())
+        return;
+
+    // totality and range
+    for (int i = 0 ; i < pattern.size() ; ++i) {
+        if (! mapping.count(i))
+            throw BuggySolution{ "No mapping for vertex " + pattern.vertex_name(i) };
+        else if (mapping.find(i)->second < 0 || mapping.find(i)->second >= target.size())
+            throw BuggySolution{ "Mapping " + pattern.vertex_name(i) + " -> " +
+                target.vertex_name(mapping.find(i)->second) + " out of range" };
+    }
+
+    // no extra stuff
+    for (auto & i : mapping)
+        if (i.first < 0 || i.first >= pattern.size())
+            throw BuggySolution{ "Vertex " + pattern.vertex_name(i.first)  + " out of range" };
+
+    // labels
+    if (pattern.has_vertex_labels())
+        for (int i = 0 ; i < pattern.size() ; ++i)
+            if (pattern.vertex_label(i) != target.vertex_label(mapping.find(i)->second))
+                throw BuggySolution{ "Mismatched vertex label for assignment " + pattern.vertex_name(i) + " -> " +
+                    target.vertex_name(mapping.find(i)->second) };
+
+    // injectivity
+    if (injective) {
+        map<int, int> seen;
+        for (auto & [ i, j ] : mapping) {
+            if (! seen.emplace(j, i).second)
+                throw BuggySolution{ "Non-injective mapping: " + pattern.vertex_name(i) + " -> " +
+                    target.vertex_name(mapping.find(i)->second) + " and " +
+                    pattern.vertex_name(seen.find(j)->second) + " -> " + target.vertex_name(j) };
+        }
+    }
+
+    // local injectivity
+    if (locally_injective) {
[…]

### 2021-01-20 glasgow-subgraph-solver/src/watches.cc

@@ -1,0 +1,4 @@
+/* vim: set sw=4 sts=4 et foldmethod=syntax : */
+
+#include "watches.hh"
+

### 2021-01-20 paper/paper.tex
Completely anonymous
Tweaks

@@
 
-% \usepackage[textsize=tiny,textwidth=2cm,shadow]{todonotes}
-% \newcommand{\skcom}[1]{\todo[color=orange!50!white]{{S}: #1}}
-% \newcommand{\rcom}[1]{\todo[color=gray!50!white]{{R}: #1}}%reviewers said
-% \usepackage{showframe}
-
 \newtheorem{proposition}{Proposition}
@@
 
-% \author{
-% Ciaran McCreesh$^1$
-% \and
-% Sonja Kraiczy$^2$
-% \affiliations
-% $^1$University of Glasgow\\
-% $^2$University of Oxford\\
-% \emails
-% ciaran.mccreesh@glasgow.ac.uk
-% }
-
 \newcommand{\neighbourhood}{\operatorname{N}}
@@
 \begin{abstract}
-    Graph homomorphism problems involve finding adjacency-preserving mappings between two graphs.
-    Many of these problems are theoretically hard, but can be solved quite efficiently in practice
-    through constraint programming techniques.  This paper adapts techniques from the
-    state-of-the-art in subgraph isomorphism solving to broader graph homomorphism problems, and
-    introduces a new form of filtering based upon clique-finding. We evaluate these techniques
-    empirically, and show they are effective in practice for the locally injective graph
-    homomorphism and subgraph isomorphism problems, as well as giving the first practical constraint
-    programming approach to finding general graph homomorphisms.
+    Graph homomorphism problems involve finding adjacency-preserving mappings between two given
+    graphs.  Although theoretically hard, these problems can often be solved in practice
+    using constraint programming algorithms. We show how techniques from the state-of-the-art in
+    subgraph isomorphism solving can be applied to broader graph homomorphism problems, and
+    introduce a new form of filtering based upon clique-finding. We demonstrate empirically that
+    this filtering is effective for the locally injective graph homomorphism and subgraph
+    isomorphism problems, and gives the first practical constraint programming approach to
+    finding general graph homomorphisms.
 \end{abstract}
@@
 problems, these applications, along with
-others in areas including compilers \cite{DBLP:conf/cp/BlindellLCS15}
+others in areas including compilers \cite{DBLP:conf/cp/BlindellLCS15},
+graph databases \cite{DBLP:journals/jair/McCreeshPST18}
 and pattern recognition \cite{DBLP:journals/ijprai/FoggiaPV14},
@@
 whose practical utility has yet to be demonstrated. The empirical effectiveness of our results is
-especially important because many current applications use subgraph isomorphism solvers only because
-they perform well using off the shelf solvers, rather than because they exactly match ideal domain
-requirements.
+especially important because many applications use subgraph isomorphism solvers only because
+they perform well off the shelf, rather than because they exactly match domain requirements.
 
@@
     Let $j$ and $k$ be two distinct vertices of the clique in $G$. By definition of a homomorphism,
-    $i(j)$ and $i(k)$ must be adjacent. And, because $H$ does not contain loops, $i(j)$ cannot equal
-    $i(k)$.
+    $i(j)$ and $i(k)$ must be adjacent. And, because $H$ has no loops, $i(j) \ne i(k)$.
 \end{proof}
@@
 has found a maximum clique. Instead, we allow it to stop as soon as it has found a clique with as
-many vertices as the largest domain clique. This is useful in practice because it may be very hard
-to decide whether a particular codomain vertex has neighbourhood clique size of, say, 15 or 16, but if
-the largest domain vertex has a neighbourhood clique size of only 5 then this is irrelevant.
+many vertices as the largest domain clique. This is useful because it may be very hard to decide
+whether a particular codomain vertex has neighbourhood clique size of, say, 15 or 16, but if the
+largest domain vertex has a neighbourhood clique size of only 5 then this is irrelevant.
 
@@
 homomorphisms, 21 additional instances; and for homomorphisms, 114, 146 or 193 additional instances
-using just distance filtering, just clique filtering, or both. Alternatively, we can measure the
-horizontal distance between plots for a given choice of timeout $t$ as follows. Let $t'$
-be the last time no greater than $t$ where the slower algorithm solved an instance, and let $u$ be
-the lowest amount of time needed for the faster algorithm to solve the same number of instances.
-The \emph{aggregate speedup}, which we plot in the top right of \cref{figure:plots}, is the ratio of
-$t'$ to $u$ or $u$ to $t'$ as appropriate. At the one hour timeout, we see aggregate speedups of 2.2
-for locally injective homomorphisms, 3.2 for subgraph isomorphism, and 47.5, 131.7, and 803.5 for
-homomorphisms.
+using just distance filtering, just clique filtering, or both. Or, at the best choices of timeouts,
+we solve 53 additional subgraph isomorphism instances; 64 additional locally injective instances;
+and 243, 383, and 424 additional homomorphism instances.
+
+Alternatively, we can measure the horizontal distance between plots for a given choice of timeout
+$t$ as follows. Let $t'$ be the last time no greater than $t$ where the slower algorithm solved an
+instance, and let $u$ be the lowest amount of time needed for the faster algorithm to solve the same
+number of instances.  The \emph{aggregate speedup}, which we plot in the top right of
+\cref{figure:plots}, is the ratio of $t'$ to $u$ or $u$ to $t'$ as appropriate. At the one hour
+timeout, we see aggregate speedups of 2.2 for locally injective homomorphisms, 3.2 for subgraph
+isomorphism, and 47.5, 131.7, and 803.5 for homomorphisms.
 
@@
 
-% \section*{Acknowledgements}
-% 
-% This work was supported by the Engineering and Physical Sciences Research Council [grant number
-% EP/P026842/1].
-
 \bibliographystyle{named}

### 2021-01-20 paper/paper.tex
Little tweaks

@@
 \begin{proposition}[neighbourhood degree sequences are preserved]\label{proposition:degreends}
-    Suppose  $i$ is a locally injective homomorphism from a $G$ to  $H$. Let
+    Suppose  $i$ is a locally injective homomorphism from $G$ to  $H$. Let
     $v$ be any vertex. Then $i$ cannot map $v$ to a vertex of lower degree, $\degree_G(v) \le \degree_H(i(v))$.
@@
 subgraph isomorphisms \cite{DBLP:journals/constraints/ZampelliDS10}; indeed, every recent constraint
-programming approach for subgraph isomorphism finding make use of degrees and neighbourhood degree
+programming approach for subgraph isomorphism finding makes use of degrees and neighbourhood degree
 sequences. The simplest way for a constraint programming algorithm to use these results is as unary

### 2021-05-20 paper/paper.tex
Revert "Completely anonymous"
This reverts commit 7b978a5c313d109c7c589c4e01ccc510dcc9999f.
Tweaks

@@
 
-\author{Paper 4726}
+\author{
+Sonja Kraiczy$^1$
+\and
+Ciaran McCreesh$^2$
+\affiliations
+$^1$University of Oxford\\
+$^2$University of Glasgow\\
+\emails
+ciaran.mccreesh@glasgow.ac.uk
+}
 
@@
 the real-world problem being solved \cite{Willett1999,EhrlichR11}. Despite being theoretically hard
-problems, these applications, along with
-others in areas including compilers \cite{DBLP:conf/cp/BlindellLCS15},
-graph databases \cite{DBLP:journals/jair/McCreeshPST18}
-and pattern recognition \cite{DBLP:journals/ijprai/FoggiaPV14},
-have given rise to a large amount of research into designing practical algorithms for solving these
-problems.  Most approaches are based either upon very fast but simple backtracking algorithms
+problems, these applications, along with others in areas including compilers
+\cite{DBLP:conf/cp/BlindellLCS15}, graph databases \cite{DBLP:journals/jair/McCreeshPST18} and
+pattern recognition \cite{DBLP:journals/ijprai/FoggiaPV14}, have given rise to a large amount of
+research into designing practical algorithms for solving these problems.  Most approaches are based
+either upon very fast but simple backtracking algorithms
 \cite{DBLP:journals/pami/CordellaFSV04,DBLP:journals/bmcbi/BonniciGPSF13,DBLP:conf/gbrpr/CarlettiFSV17}
@@
 \cite{DBLP:journals/dam/FialaKK01,DBLP:conf/semweb/Baget05,DBLP:conf/iccs/CorbyF07,DBLP:journals/csr/FialaK08,DBLP:journals/pvldb/FanLMWW10,DBLP:journals/tcs/ChaplickFHPT15}.
-(There is also an edge-surjective version of the problem known as an \emph{epimorphism}
-\cite{DBLP:journals/dam/GayFMSS14}.)
 
@@
 We begin by introducing notation and terminology, and by giving the theory which
-support our implementation.
+supports our implementation.
 
@@
     \begin{enumerate}
-        \item The distance between $v$ and $w$ is at least the distance between $i(v)$ and $i(w)$
+        \item The distance between $v$ and $w$ is at most the distance between $i(v)$ and $i(w)$
             \cite{DBLP:conf/cp/AudemardLMGP14}.
@@
 prevents any kind of merging, local injectivity only prevents two vertices from being merged if they
-are both in the neighbourhood of the same vertex. However, we observe that a weaker result does
-hold:
+are both in the neighbourhood of the same vertex. However, a weaker result does hold:
 
@@
 isomorphism between several different pairs of graphs. Let $G^d$ be the graph with the same set of
-vertices as $G$, but an edge between vertices $v$ and $w$ if the distance between $v$ and $w$ in $G$
-is at most $d$.  Similarly, let $G^{n,\ell}$ be the graph with the same set of vertices as $G$, but
-an edge between vertices $v$ and $w$ if there are at least $n$ simple paths of length exactly $\ell$
-between $v$ and $w$ in $G$. Following on from \cref{proposition:distances,proposition:paths}, we
-generalise the result of \citeauthor{DBLP:conf/cp/McCreeshP15} as follows.
+vertices as $G$, but with an edge between vertices $v$ and $w$ if the distance between $v$ and $w$
+in $G$ is at most $d$.  Similarly, let $G^{n,\ell}$ be the graph with the same set of vertices as
+$G$, but an edge between vertices $v$ and $w$ if there are at least $n$ simple paths of length
+exactly $\ell$ between $v$ and $w$ in $G$. Following on from
+\cref{proposition:distances,proposition:paths}, we generalise the result of
+\citeauthor{DBLP:conf/cp/McCreeshP15} as follows.
 
@@
 
-Given the complexity of modern solvers, a critical question is how we can be sure
-that they are producing correct answers. The Glasgow Subgraph Solver's subgraph isomorphism and
-clique algorithms are both \emph{certifying} \cite{DBLP:journals/csr/McConnellMNS11}: that is, they
-can output a mathematical proof that they have reached a correct answer by sound reasoning, and this
-proof log can be verified by the VeriPB proof checking tool
-\cite{DBLP:conf/cp/GochtMMNPT20,DBLP:conf/ijcai/GochtMN20}.
-
-Implementing proof-logging for the changes not involving clique-finding was routine---although proof
-logging did catch an implementation bug where we were incorrectly applying degree-based filtering on
-the $G^{2,n}$ graphs for locally injective homomorphisms, in violation of the counter-example to
-\cref{corollary:lishapes}. For clique-based filtering, the process involved substantially more
-technical work. Briefly, for a given domain vertex $v$ that cannot be mapped to a codomain vertex
-$w$, it is possible to instruct the proof verifier to derive a set of conditional clique constraints
-saying that ``either $v$ cannot be mapped to $w$, or the following clique-like problem must be
-satisfiable'' by following steps similar to the maximum common subgraph to clique reduction
-presented by \citet{DBLP:conf/cp/GochtMMNPT20}, and then to reuse clique
-proof-logging techniques to show that this clique-like problem is unsatisfiable.
+Given the complexity of modern solvers, a critical question is how we can be sure that they are
+producing correct answers. The Glasgow Subgraph Solver's subgraph isomorphism and clique algorithms
+are both \emph{certifying} \cite{DBLP:journals/csr/McConnellMNS11}: that is, they can output a
+mathematical proof that they have reached a correct answer by sound reasoning, and this proof log
+can be verified by the VeriPB proof checking tool
+\cite{DBLP:conf/cp/GochtMMNPT20,DBLP:conf/ijcai/GochtMN20}. We therefore extended the solver's
+existing proof logging support to cover our additions. For the changes not involving clique-finding,
+this was entirely routine---although proof logging did catch an implementation bug where we were
+incorrectly applying degree-based filtering on the $G^{2,n}$ graphs for locally injective
+homomorphisms, in violation of the counter-example to \cref{corollary:lishapes}. For clique-based
+filtering, the process involved substantially more technical work. Briefly, for a given domain
+vertex $v$ that cannot be mapped to a codomain vertex $w$, it is possible to instruct the proof
+verifier to derive a set of conditional clique constraints saying that ``either $v$ cannot be mapped
+to $w$, or the following clique-like problem must be satisfiable'' by following steps similar to the
+maximum common subgraph to clique reduction presented by \citet{DBLP:conf/cp/GochtMMNPT20}, and then
+to reuse clique proof-logging techniques to show that this clique-like problem is unsatisfiable.  We
+were thus able to produce and verifying proofs of correctness for medium-sized instances, giving us
+confidence our implementation was correct.
 
 This situation is not entirely ideal. The proofs produced this way are inherently dependent upon a
-\emph{particular} choice of domain and codomain vertices. When not proof logging, we only need to
-solve a single maximum clique problem for each vertex in the domain, and a restricted clique problem
-for each vertex in the codomain. With proof logging, we must instead solve restricted clique
-problems for each assignment that is to be filtered, which is potentially a linear factor blowup.
-This seems wasteful, since for any given codomain vertex these proofs are effectively the same up to
-substitution of variable names. It would therefore seem desirable to extend the proof system with a
-substitution rule which would allow such proofs to be recycled to avoid duplication.
+\emph{particular} choice of domain and codomain vertices. Thus, when proof logging, we must solve
+restricted clique problems for each assignment that is to be filtered using a clique rule, rather
+than for each clique found; this is potentially a linear factor blowup.  This seems wasteful, since
+for any given codomain vertex these proofs are effectively the same up to substitution of variable
+names. It would therefore seem desirable to extend the proof system with a substitution rule which
+would allow such proofs to be recycled to avoid duplication.
 
@@
 with dual Intel Xeon E5-2697A v4 CPUs with 512GBytes of RAM, running Ubuntu 18.04, and using a
-timeout of one hour. Because injectivity relaxations have received less attention until now, for all
-three problem variants we will use a suite of 14,621 benchmark instances designed for the subgraph
-isomorphism problem \cite{DBLP:conf/lion/KotthoffMS16}. Note that 8,904 of these instances contain
-at least one loop in the codomain graph, and so are trivial for the homomorphism problem---we still
-include these to avoid using different y-axes on different plots.
+timeout of one hour; we do not enable proof logging for these benchmarks since proof logging
+introduces large slowdowns due to I/O costs \cite{DBLP:conf/cp/GochtMMNPT20}. Because injectivity
+relaxations have received less attention until now, for all three problem variants we will use a
+suite of 14,621 benchmark instances designed for the subgraph isomorphism problem
+\cite{DBLP:conf/lion/KotthoffMS16}. Note that 8,904 of these instances contain at least one loop in
+the codomain graph, and so are trivial for the homomorphism problem---we still include these to
+avoid using different y-axes on different plots.
 
@@
 
+\section*{Acknowledgements}
+
+This work was supported by the Engineering and Physical Sciences Research Council [grant number
+EP/P026842/1].
+
 \bibliographystyle{named}

### 2021-05-27 paper/paper.tex
Editorial tweaks

@@
 
-\paragraph{Degree-based invariants.} Constraint programming techniques for finding
-homomorphisms can be made much more effective by exploiting certain invariants. It is
-straightforward to verify the following.
+\subsection{Invariants}
+
+Constraint programming techniques for finding homomorphisms can be made much more effective by
+exploiting certain invariants. It is straightforward to verify the following.
 
@@
 neighbourhood degree sequence is too low may immediately be rejected.
+However, these invariants do \emph{not} hold for homomorphisms in general. As a
+counter-example, a homomorphism may map a star graph onto a single edge, as in
+\cref{figure:counterexamples} (top left).
 
@@
 
-However, these invariants do \emph{not} hold for homomorphisms in general. As a
-counter-example, a homomorphism may map a star graph onto a single edge, as in
-\cref{figure:counterexamples} (top left).
-
-\paragraph{Path-based invariants.} Constraint programming solvers can also exploit
-invariants that are based upon paths. This is done by automatically adding \emph{implied}
-constraints to the problem that are implied by the original model, but that will give stronger
-propagation power.
+Constraint programming solvers can also exploit invariants that are based upon paths. This is done
+by automatically adding \emph{implied} constraints to the problem that are implied by the original
+model, but that will give stronger propagation.
 
@@
 
-\paragraph{Cliques and homomorphisms.} We now discuss the main new technique of this paper. Observe
+We now discuss the main new technique of this paper. Observe
 that so far, we have not seen any unary constraints which are valid for the general homomorphism
@@
 to reuse clique proof-logging techniques to show that this clique-like problem is unsatisfiable.  We
-were thus able to produce and verifying proofs of correctness for medium-sized instances, giving us
-confidence our implementation was correct.
+were able to produce and verify proofs of correctness for medium-sized instances, giving us
+confidence in our implementation.
 
